### 2025-04-08 01_tidying_data.qmd
Update before optimising

---
title: "Tidying data"
format: html
editor: visual
---

Setup

```{r}

library(tidyverse)
library(here)
library(terra)
library(qs)

#model functions
source("src/dir.R")
source("src/model_functions.R")

#species of interest

species <- c("atlantic_salmon", 
             "general_salmonid", 
             "european_seabass", 
             "gilthead_seabream", 
             "large_yellow_croaker", 
             "japanese_seabass",
             "japanese_amberjack",
             "general_grouper",
             "pompano",
             "milkfish")

source(here("src/spatial_templates.R"))

```

Make repo structure

```{r}

create_repo(this_path = here())

```

Make input and output folder structure for species of interest

```{r}

#deploy data folder structure for species

map(.x = species, .f = create_species_folders)

```

Pull in mariculture locations data from Clawson et al 2022 and save the species groups of interest

```{r}

farm_locations <- read.csv(file.path(rdsi_dir, "raw_data/aquaculture-locations/all_marine_aquaculture_farms_sources_final.csv")) |> 
  filter(species_group %in% c("Salmonidae fish", "General marine fish"))

qsave(x = farm_locations, file = "data/_general_data/farm_locations/locations.qs")

```

Modify Feed characterisation and save to all species forcings

```{r}

#import food characterisation data - this is calculated in an independent google drive document

diet = "reference"

Food_charactisation_list <- 
  read_csv(sprintf("data/_general_data/diets/protein_lipid_carb_summaries_%s.csv", diet)) |> 
  select(-c(5)) |> 
  pivot_longer(names_to = "nutrient", values_to = "prop", cols = -species) |> 
  group_by(species) |> 
  group_split()

map(.x = Food_charactisation_list, .f = \(this_species_food){

  this_species <- this_species_food$species |> unique()

  this_species_food_df <- 
    this_species_food |> 
    relocate(prop, .before = "nutrient") |> 
    mutate(nutrient = as.character(paste0(str_to_sentence(nutrient), "s"))) |> 
    mutate(prop = as.double(prop)) |> 
    select(-species)

  write.table(this_species_food_df, file = sprintf("data/%s/forcings/Food_characterization_%s.csv", this_species, diet), row.names = FALSE, col.names = FALSE, sep = ",")

})

#now repeat for future feed

diet = "future"

Food_charactisation_list <- 
  read_csv(sprintf("data/_general_data/diets/protein_lipid_carb_summaries_%s.csv", diet)) |> 
  pivot_longer(names_to = "nutrient", values_to = "prop", cols = -species) |> 
  group_by(species) |> 
  group_split()

map(.x = Food_charactisation_list, .f = \(this_species_food){

  this_species <- this_species_food$species |> unique()

  this_species_food_df <- 
    this_species_food |> 
    relocate(prop, .before = "nutrient") |> 
    mutate(nutrient = as.character(paste0(str_to_sentence(nutrient), "s"))) |> 
    mutate(prop = as.double(prop)) |> 
    select(-species)

  write.table(this_species_food_df, file = sprintf("data/%s/forcings/Food_characterization_%s.csv", this_species, diet), row.names = FALSE, col.names = FALSE, sep = ",")

})

#now repeat for past feed

diet = "past"

Food_charactisation_list <- 
  read_csv(sprintf("data/_general_data/diets/protein_lipid_carb_summaries_%s.csv", diet)) |> 
  pivot_longer(names_to = "nutrient", values_to = "prop", cols = -species) |> 
  group_by(species) |> 
  group_split()

map(.x = Food_charactisation_list, .f = \(this_species_food){

  this_species <- this_species_food$species |> unique()

  this_species_food_df <- 
    this_species_food |> 
    relocate(prop, .before = "nutrient") |> 
    mutate(nutrient = as.character(paste0(str_to_sentence(nutrient), "s"))) |> 
    mutate(prop = as.double(prop)) |> 
    select(-species)

  write.table(this_species_food_df, file = sprintf("data/%s/forcings/Food_characterization_%s.csv", this_species, diet), row.names = FALSE, col.names = FALSE, sep = ",")

})

```

Daily SST data from NASA - GHRSST Level 4 MUR 0.25deg Global Foundation Sea Surface Temperature Analysis

```{r}

nasa_dir <- "/mnt/rdsi/raw_data/NASA/GHRSST_L4_MUR_0.25_SST/"

#rename the NASA files

#SST_files <- list.files(file.path(rdsi_dir, "raw_data/NASA/GHRSST_L4_MUR_0.25_SST"), full.names = FALSE, pattern = ".nc")

# map(SST_files, \(this_file){
#   file.rename(from = file.path(nasa_dir, this_file), to = file.path(nasa_dir, sub('^(.{4})(.*)$', '\\1_\\2',  this_file)))
# })

SST_files <- list.files(file.path(rdsi_dir, "raw_data/NASA/GHRSST_L4_MUR_0.25_SST"), full.names = TRUE, pattern = ".nc")

# this_file <- SST_files[[1]]
# 
# this_year=2013

#initialise lists for each year

years <- c(2010:2019)
year_lists <- vector(mode ="list", length = length(years))

# Add annual files to a list of files for each year

for(y in 1:length(years)){

  this_year <- years[y]

  this_years_files <- SST_files[grep(paste0(this_year, "_"), SST_files)]

  this_years_files <- this_years_files[!grepl(paste0(this_year, "_0229"), this_years_files)]

  year_lists[[y]] <- this_years_files

}

# Extract the daily files for each year and rasterize in a stack

days <- c(1:365)

this_day <- days[1]

rast_list = vector(mode = "list", length = length(days))

map(days, \(this_day){

  this_days_file_list <- sapply(year_lists, "[[", this_day)

  message("SST layer for Day ", this_day)

  this_day_stack <- rast(this_days_file_list, subds = "analysed_sst")

  this_day_mean_rast <- app(this_day_stack, "mean", na.rm=TRUE)-273.15

  writeRaster(this_day_mean_rast, filename = sprintf(here("data/_general_data/SST/SST_rasters/sst_nasa_mur_L4_0.25_mean2010-2019_day_%s.tif"), this_day))

})

#now gapfill the saved rasters

raw_SST_files <- list.files("data/_general_data/SST/SST_rasters/", full.names = TRUE)

#test function
#this_file <- raw_SST_files[[1]]

map(.x = raw_SST_files, .f = \(this_file){

  saveName <-  basename(this_file)

  message("Gapfilling ", saveName)

  this_rast <- rast(this_file)

  this_gf_rast <- focal(this_rast, w=7, fun = "mean", na.policy = "only")

  writeRaster(this_gf_rast, filename = sprintf("data/_general_data/SST/SST_gf_rasters/%s", saveName))

})

```

Pull in spp vulnerability to eutrophication data - save to salmon folder

```{r}

vuln_scores <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_score.csv")) |> 
  select(vuln_gf_id, eutrophication_nutrient_pollution)

vuln_sd <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_sd.csv")) |> 
  select(vuln_gf_id, eutrophication_nutrient_pollution) |> rename(sd=eutrophication_nutrient_pollution)

vuln_tx <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_tx.csv"))

full_vuln_df <- 
  vuln_scores |> 
  left_join(vuln_sd) |> 
  left_join(vuln_tx)

qsave(x = full_vuln_df, file = here("data/_general_data/species_layers/vulnerabilities/marine_spp_vulnerabilities_eutrophication.qs"))

```

Pull in spp distribution data from Aquamaps

```{r}

aquamaps_0.5d <- data.table::fread(file.path(rdsi_data_dir, "aquamaps/aquamaps_0.6_depth_prepped.csv"))

#all cells have an id. These id's can be used to isolate only the cells of relevance for the salmon farming locations.
filter(aquamaps_0.5d, is.na(cell_id))

#save to project
qsave(x = aquamaps_0.5d, file = here("data/_general_data/species_layers/distribution/aquamaps_0.5d.qs"))

```