### 2025-04-08 07_methods_and_supplementaries/07_methods_and_supplementaries.qmd

---
title: "Model inner workings - for methods and supplementary sections"
author: Tormey Reimer
date: last-modified
bibliography: ["../resources/refs.bib", "../resources/manual-refs.bib"]
format: 
  html:
    theme: simplex
    css: ["../resources/extra.css"]
    code-fold: true
    code-overflow: wrap
    toc: true
    toc-expand: true
    toc-location: left
    lang: en-GB
    grid:
      sidebar-width: 350px
      body-width: 1200px
      margin-width: 250px
      gutter-width: 2em
  # pdf:
  #   papersize: a4
  #   linestretch: 1.15
  #   link-citations: true
  #   keep-tex: true
  #   template-partials:
  #     - "C:/Users/treimer/Documents/R-temp-files/local_to_global_mariculture_modelling/resources/title.tex"
  #   include-in-header:
  #     file: "C:/Users/treimer/Documents/R-temp-files/local_to_global_mariculture_modelling/resources/header_extra.tex"

filters:
  - acronyms

execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true

knitr: 
  opts_chunk:
    fig.align: center

editor: source
---

# Introduction

```{r setup}
#| include: false
#| cache: false
#| eval: true

packs <- unique(c('magrittr', 'dplyr', 'tidyr', 'stringr', 'ggplot2', 'here', 'ggpubr', 'patchwork', 'qs', 'arrow', 'targets', 'sf', 'terra', 'rnaturalearth', 'rnaturalearthdata', 'dtplyr', 'furrr', 'RColorBrewer', 'conflicted'))
pacman::p_load(char = packs)

conflicts_prefer(dplyr::select(), dplyr::filter(), .quiet = T)

source(here("src/spatial_templates.R"))
source(here("src/model_functions.R"))

# Sys.setenv(TAR_PROJECT = "project_main")
base_path <- file.path("C:", "Users", "treimer", "Documents", "R-temp-files", "local_to_global_mariculture_modelling")
targets_store <- file.path(base_path, "targets", "_targets_individual")
farm_IDs <- tar_read(farm_IDs, store = targets_store)
b1 <- sample(length(farm_IDs), 100, replace = F)
b2 <- b1 + length(farm_IDs)
b3 <- b2 + length(farm_IDs)
b <- c(b1, b2, b3)
```

# Temporary questions to answer:

-   Do the fish reach harvest size within a reasonable amount of time?
    -   If not, are they growing for the correct amount of time, starting at the correct weight?
-   Is their FCE/FCR reasonably close to experimental data?
-   Is their SGR reasonably close to experimental data?

# Model approach

# Water temp

```{r water-temp-data}
#| code-summary: Load and cache data to avoid spending a long time on it every render.

# This is the full idealised year (2 years?) of temperature data
sst <- file.path(base_path, "data", "_general_data", "SST", "farm_SST_extracted.parquet") %>% read_parquet()

# This is the actual temperature curve for the farming period
farm_temp <- tar_read(example_individual, store = targets_store, branches = b1)
```

Originally, all salmon were transferred to grow-out cages (model began) on January 1st. This isn't particularly realistic. Now, all farms begin the modelling period in spring (1st of May in the northern hemisphere, 1st of October in the southern hemisphere).

```{r fig-water-temp}
#| fig-cap: Water temperature for the production period, starting at either DOY 121 (1st May, northern hemisphere) or DOY 274 (1st October, southern hemisphere).

farm_temp %>% 
  ggplot(aes(x = prod_days, y = water_temp, colour = as.factor(farm_ID))) +
  geom_line(linewidth = 0.75) +
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 550, 50)) +
  labs(x = "Day of production cycle", y = expression("Water temperature ("*degree*"C)"))
```

# Weight

```{r weight-related-data}
#| code-summary: Load and cache data to avoid spending a long time on it every render.

species_params <- tar_read(species_params, store = targets_store)
pop_params <- tar_read(pop_params, store = targets_store)
weight <- tar_read(example_individual, store = targets_store, branches = b) %>% select(farm_ID, weight, days, prod_days, feed)
SGR <- tar_read(example_individual, store = targets_store, branches = b) %>% select(farm_ID, SGR, days, prod_days, feed)
FCR <- tar_read(example_individual, store = targets_store, branches = b) %>% select(farm_ID, FCR, days, prod_days, feed)

weight_smolt <- weight %>% 
  filter(prod_days == 152 & feed == "reference") %>% 
  group_by(feed) %>% 
  reframe(final_weight = mean(weight))
weight_ocean <- weight %>% 
  filter(prod_days == max(prod_days) & feed == "reference") %>% 
  group_by(feed) %>% 
  reframe(final_weight = mean(weight),
          final_wt_sd = sd(weight))
```

@fig-ref-weight shows the change in weight for 100 individuals grown at different farms. Within the first 5 months (post-smolt period) the fish grow from `r unname(pop_params['meanW'])`g to `r round(weight_smolt$final_weight, 1)`g, or approximately `r round(weight_smolt$final_weight/125, 1)`$\times$ their starting weight. By the end of the production cycle (730 days) the fish have grown to a mean of `r round(weight_ocean$final_weight, 1)` g with a SD of `r round(weight_ocean$final_wt_sd, 1)` g. 
<!-- This is not quite what's needed - I'm expecting individual weights to at least approximate the mean commercial weight of 5kg. -->

```{r fig-ref-weight}
#| fig-cap: Weight of single individuals grown at 100 random farms across the production period, starting at either DOY 121 (1st May, northern hemisphere) or DOY 274 (1st October, southern hemisphere).

weight %>% 
  mutate(feed == "reference") %>% 
  ggplot(aes(x = prod_days, y = weight, colour = as.factor(farm_ID))) +
  geom_line(linewidth = 0.75) +
  scale_y_continuous(breaks = seq(0, 7500, 500)) +
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 550, 150)) +
  labs(x = "Day of production cycle", y = "Individual weight (g)")
```

```{r fig-ref-SGR}
#| fig-cap: Specific growth rate of single individuals grown at 100 random farms across the production period, starting at either DOY 121 (1st May, northern hemisphere) or DOY 274 (1st October, southern hemisphere).

SGR %>% 
  mutate(feed == "reference") %>% 
  ggplot(aes(x = prod_days, y = SGR, colour = as.factor(farm_ID))) +
  geom_line(linewidth = 0.75) +
  scale_y_continuous(breaks = seq(0, 2.5, 0.25)) +
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 550, 50)) +
  labs(x = "Day of production cycle", y = expression("SGR"))
```

```{r fig-ref-FCR}
#| fig-cap: FCR of 25 random farms across the production period, starting at either DOY 121 (1st May) or DOY 274 (1st October).

FCR %>% 
  mutate(feed == "reference") %>% 
  filter(FCR < 50 & FCR > -50) %>% 
  group_by(prod_days) %>% 
  reframe(sd = sd(FCR),
          FCR = mean(FCR)) %>% 
  ggplot(aes(x = prod_days, y = FCR, ymin = FCR-sd, ymax = FCR+sd)) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25) +
  theme_classic() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 550, 50)) +
  labs(x = "Day of production cycle", y = expression("FCR"))
```

# General fish functions

@fig-functional-response-to-temperature shows the metabolic response of all salmon to temperature (affecting their relative metabolism), and @fig-feeding-rate-with-temperature shows how the salmons' feeding rate changes with temperature.

```{r fig-functional-response-to-temperature}
#| fig-cap: Metabolic response of all salmon to temperature within the model.

df <- data.frame(water_temp = seq(-5, 30, 0.25))
df$resp <- exp(species_params['pk']*df$water_temp)

df %>% 
  ggplot(aes(x = water_temp, y = resp)) +
  geom_line(linewidth = 0.75) +
  scale_y_continuous(breaks = seq(0,5,0.5)) +
  scale_x_continuous(breaks = seq(-10,35,5)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Temperature response") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

$$
cat = \epsilon_{O_2} \times k_0 \times T_{resp} \times W^n \times \omega
$$

Relative feeding rate is temperature-dependent and calculated via:

$$
FR_{rel} = e^{b(T_w-T_{opt})} \times \bigg[\frac{T_{max}-T_w}{T_{max}-T_{opt}}\bigg]^{b(T_{max}-T_{opt})}
$$

where $T_{opt}$ is the optimum feeding temperature, $T_{max}$ is the maximum feeding temperature (lethal temperature), $T_w$ is the current water temperature, and $b$ is a species-specific shape coefficient.

```{r fig-feeding-rate-with-temperature}
#| fig-cap: Changes in salmons feeding rate with temperature. The dashed line shows the optimum feeding temperature for salmon while the dotted lines show the minimum and maximum feeding temperatures.

df <- data.frame(water_temp = seq(-5, 30, 0.25))
df$feeding <- sapply(FUN = feeding_rate, X = df$water_temp, species_params = species_params)

df %>% 
  ggplot(aes(x = water_temp, y = feeding)) +
  geom_line(linewidth = 0.75) +
  geom_vline(aes(xintercept = species_params["Toa"]), linetype = "dashed") +
  geom_vline(aes(xintercept = species_params["Taa"]), linetype = "dotted") +
  geom_vline(aes(xintercept = species_params["Tma"]), linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative feeding rate") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

# Feed data

```{r feed-raw-data}
#| code-summary: Load and cache data to avoid spending a long time on it every render.

feed_params_protein <- tar_read(feed_params_protein, store = targets_store)
feed_params_carbs <- tar_read(feed_params_carbs, store = targets_store)
feed_params_lipids <- tar_read(feed_params_lipids, store = targets_store)
```

Incorporated the individual digestibility of each ingredient and switched to tracking ingredients separately instead of feed – this unfortunately makes the model run slower but I think it will be worth it once the digestibility coefficients from the experiments are incorporated.

```{r feed-data}

ref_feed <- list(
    feed_params_protein[[1]] %>% 
      mutate(feed = "reference", type = "protein") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_carbs[[1]] %>% 
      mutate(feed = "reference", type = "carb") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_lipids[[1]] %>% 
      mutate(feed = "reference", type = "lipid") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest))
) %>% bind_rows()
past_feed <- list(
    feed_params_protein[[2]] %>% 
      mutate(feed = "past", type = "protein") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_carbs[[2]] %>% 
      mutate(feed = "past", type = "carb") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_lipids[[2]] %>% 
      mutate(feed = "past", type = "lipid") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest))
) %>% bind_rows()
fut_feed <- list(
    feed_params_protein[[3]] %>% 
      mutate(feed = "future", type = "protein") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_carbs[[3]] %>% 
      mutate(feed = "future", type = "carb") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest)),
  feed_params_lipids[[3]] %>% 
      mutate(feed = "future", type = "lipid") %>% group_by(feed, type) %>% 
      reframe(proportion = sum(proportion),
              macro = sum(macro),
              digest = sum(digest))
) %>% bind_rows()

feeds <- rbind(ref_feed, past_feed, fut_feed)
```

```{r fig-compare-feeds-composition}
#| fig-cap: Feed composition

```

```{r fig-compare-feeds-digestibility}

```

# Individual runs

I set up some “example fish” to speed up future model adjustments – basically fish that are the average of their whole farm, easier than running 5000 fish per farm while I’m making changes.

## Food provided vs food eaten

```{r food-raw-data}
#| code-summary: Load and cache data to avoid spending a long time on it every render.

#c(b1, b2, b3))
ing_pot <- tar_read(example_individual, store = targets_store, branches = b1) %>% select(farm_ID, ing_pot, days, prod_days, feed)
ing_act <- tar_read(example_individual, store = targets_store, branches = b1) %>% select(farm_ID, ing_act, days, prod_days, feed)
food_prov <- tar_read(example_individual, store = targets_store, branches = b1) %>% select(farm_ID, food_prov, days, prod_days, feed)
```

Within the model, salmon have a maximum ingestion potential (based on their weight and individualised feeding rate). The actual food ingested is 97% of their ingestion potential (food encounter efficiency) or the total food provided, whichever is less. @fig-food-prov-theoretical shows an example of how food provided scales with potential individual ingestion.

```{r fig-food-prov-theoretical}
#| fig-cap: Example of food provided based on potential ingestion (where feeding rate ranges from 0 to 1). This curve is constructed with a single fish of mean starting weight (125 g) and average maximum ingestion rate (0.035). The dashed line shows food provision == potential ingestion.

df <- data.frame(rel_feeding = seq(0,1,0.01)) %>% 
mutate(ing_pot = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * rel_feeding,
       food_prov = NA)
for (i in 1:nrow(df)) {
  df$food_prov[i] <- food_prov_rate(rel_feeding = df$rel_feeding[i], ing_pot = df$ing_pot[i], 
                                    ing_pot_10 = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * 0.1)
}
lin <- seq(0,max(df$ing_pot),length.out=101)

df %>% 
  ggplot(aes(x = ing_pot, y = food_prov)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(x = lin, y = lin), linetype = "dashed") +
  theme_classic() +
  scale_x_continuous(breaks = seq(0,1.5, 0.25), limits = c(0,max(df$food_prov))) +
  scale_y_continuous(breaks = seq(0,1.5, 0.25)) +
  labs(x = "Potential ingestion (g)", y = "Food provided (g)")
```

```{r feed-nums}

prov <- merge(ing_pot, ing_act, by = c("farm_ID", "feed", "days", "prod_days")) %>% 
  merge(food_prov, by = c("farm_ID", "feed", "days", "prod_days")) %>% 
  mutate(uneaten = food_prov - ing_act,
         uneaten_perc = uneaten/food_prov) %>% 
  select(-c(food_prov, ing_act))

# Therefore, uneaten feed can be quite high (up to \~30%) when relative feeding is $\leq$10%. But generally, the median amount of uneaten food is `r round(100*median(prov$uneaten_perc), 1)`% (@fig-uneaten-feed).
```

```{r fig-uneaten-feed}
#| fig-cap: Difference between food provided and actual ingestion (i.e. total food ingestion efficiency) for 100 random farms.

prov %>% 
  ggplot(aes(x = ing_pot, y = 100*uneaten_perc, colour = feed)) +
  geom_point() +
  theme_classic() +
  scale_colour_brewer(palette = "Set1") +
  theme_classic() +
  labs(x = "Potential feed ingestion (g)", y = "Uneaten food (% of provided)")
```

# Parameters for Atlantic salmon

```{r salmon params}
#| code-summary: lsdkjfk
#| eval: true

params <- tar_read(sens_all_params, store = targets_store)
pnames <- names(params)
```

| Parameter | Value | Units | Description | Reference | Comment | 
|:-------:|:----:|:----:|-------------------|-------------|---------------------------|
| $\alpha$ | `r unname(params['alpha'])` | - | Feeding catabolism coefficient | @broekhuizen_modelling_1994 | *Salvelinus alpinus*, *Oncorhynchus mykiss* (temps: 8–13$^{\circ}$C) | 
| $\epsilon_{P}$ | `r unname(params['epsprot'])` | J g protein-1 | Energy content of protein | @baldan_r_2018 | | 
| $\epsilon_{L}$ | `r unname(params['epslip'])` | J g lipid-1 | Energy content of lipid | @baldan_r_2018 | | 
| $\epsilon_{C}$ | `r unname(params['epscarb'])` | J g carbohydrate-1 | Energy content of carbohydrate | @baldan_r_2018 | | 
| $\epsilon_{O_2}$ | `r unname(params['epsO2'])` | J g O2-1 | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 | | 
| $pk$ | `r unname(params['pk'])` | d-1 | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 | | 
| $k_0$ | `r unname(params['k0'])` | degree C-1 | Fasting catabolism at 0 Celsius degree | @macnaughton_using_2019 | | 
| $m$ | `r unname(params['m'])` | - | Weight exponent for the anabolism | @broekhuizen_modelling_1994 | *Salmo trutta* (temps: 8–13$^{\circ}$C) | 
| $n$ | `r unname(params['n'])` | - | Weight exponent for the catabolism | @baldan_r_2018 | | 
| $\beta C$ | `r unname(params['betac'])` | - | Shape coefficient for the feeding function | @jonsson_thermal_2001 | Fish size range: 2.25$\pm$0.49 – 11.5$\pm$5.29 g, temperature ranges: 5–24$\^{\circ}$C. Coefficent for mean daily energy intake of the moderately fast growing group, 0.4398 (mean fast group), 0.2214 (mean slow group) | 
| $T_{max}$ | `r unname(params['Tma'])` | C | Maximum lethal temperature | @poletto_unusual_2017 | Juvenile Chinook salmon (19.8$\pm$0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) | 
| $T_{opt}$ | `r unname(params['Toa'])` | C | Optimal temperature | @jonsson_thermal_2001 | Fish size range: 2.25$\pm$0.49 – 11.5$\pm$5.29 g, temperature ranges: 5–24$^{^\circ}$C.
Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0$\pm$14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$ | `r unname(params['Taa'])` | C | Lowest feeding temperature | @vadboncoeur_lowering_2023 | Fish essentially stop feeding when below 2$^{\circ}$C | | $\omega$ | `r unname(params['omega'])` | g O2 / g | Oxygen consumption:weight loss ratio | @baldan_r_2018 | Calibrated value | 
| $a$ | `r unname(params['a'])` | J g tissue-1 | Energy content of fish tissue | @van_tien_development_2016 | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 | 
| $k$ | `r unname(params['k'])` | - | Weight exponent for energy content | @van_tien_development_2016 | Alt value: 1.65175 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 | 
| $eff$ | `r unname(params['eff'])` | - | Food ingestion efficiency | @uglem_does_2020 | Range: 95-97% ingestion efficiency | 
| $\overline{W}$ | `r unname(params['meanW'])` | g | Dry weight mean | @dempsey_estimating_2023 | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers | 
| $\Delta{W}$ | `r unname(params['deltaW'])` | g | Dry weight standard deviation | @dempsey_estimating_2023 | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers | 
| $Wlb$ | `r unname(params['Wlb'])` | g | Dry weight lower bound | Reference | | 
| $\overline{I_{max}}$ | `r unname(params['meanImax'])` | g g-1 fish day-1 | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) | | 
| $\Delta{I_{max}}$ | `r unname(params['deltaImax'])` | g g-1 fish day-1 | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) | | 
| $mort$ | `r unname(params['mortmyt'])` | d-1 | Natural mortality rate | @tvete_towards_2023 | |
: Params

## Sensitivity

In the following plots, a value > 0 indicates that increasing that parameter increases the output variable while a negative value indicates that increasing the parameter decreases the variable. The sensitivity follows a power relationship, such that output$_1$ $\propto$ output$_0^{S}$.

```{r param sensitivity data}
#| code-summary: Load sensitivity data

sens_wt <- file.path("..", "data", "atlantic_salmon", "data_products", "weight_parameter_sensitivity.parquet") %>% 
  read_parquet() %>% 
  mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
# sens_dw <- file.path("data", "atlantic_salmon", "data_products", "dw_parameter_sensitivity.parquet") %>% 
#   read_parquet() %>% 
#   mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
# sens_excr <- file.path("data", "atlantic_salmon", "data_products", "excreted_parameter_sensitivity.parquet") %>% 
#   read_parquet() %>% 
#   mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
# sens_uneat <- file.path("data", "atlantic_salmon", "data_products", "uneaten_parameter_sensitivity.parquet") %>% 
#   read_parquet() %>% 
#   mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
```

Looking at @fig-param-sens-weight:

* The most sensitive parameter by far is $m$ the weight exponent for anabolism, followed closely by $n$ the weight exponent for catabolism
* These two are also relatively uncertain, as $m$ was taken from *Salmo trutta* measurements in a fairly old paper [@broekhuizen_modelling_1994] and $n$ was taken from @baldan_r_2018 which did not specifically look at Atlantic salmon. There must be more up-to-date measurements for Atlantic salmon in particular.

```{r fig-param-sens-weight}
#| fig-cap: Impact of varying each model parameter from 0.9-1.1 on individual fish weight at the end of the production period.

(
  p_sens <- sens_wt %>% 
    ggplot(aes(x = adj_param, y = sens, ymin = sens-sd, ymax = sens+sd)) +
    geom_col(alpha = 0.35, colour = "black") +
    geom_errorbar(width = 0.3) +
    geom_hline(aes(yintercept = 0), linetype = "dashed") +
    scale_y_continuous(breaks = seq(-20,30,5)) +
    labs(x = "Adjusted parameter", y = "Sensitivity") +
    theme_classic() + theme(legend.position = "none", 
          strip.text = element_blank(), 
          aspect.ratio = 0.8,
          text = element_text(size = 12)) +
    coord_flip()
 )
```

```{r fig-param-sens-dw}
#| fig-cap: Impact of varying each model parameter from 0.9-1.1 on the mean daily weight change of individual fish across the entire production period.
#| eval: false

p_sens %+% sens_dw
```

```{r fig-param-sens-excr}
#| fig-cap: Impact of varying each model parameter from 0.9-1.1 on the total excretion (g protein, carbs and lipids) of an individual fish across the entire production period.
#| eval: false

p_sens %+% sens_excr
```

```{r fig-param-sens-uneat}
#| fig-cap: Impact of varying each model parameter from 0.9-1.1 on the total uneaten feed (g) of an individual fish across the entire production period.
#| eval: false

p_sens %+% sens_uneat
```

# References