### 2025-06-04 manuscript.qmd
Moving text to Google doc

---
title: "Local to global N input modelling"
# Author roles: "writing – original draft", "writing – review & editing", "formal analysis", "funding acquisition", "conceptualization", "data curation", "investigation", "methodology", "project administration", "resources", "software", "supervision", "validation", "visualization"

bibliography: ["refs.bib", "manual-refs.bib"]

format: 
  html:
    theme: simplex
    css: ["extra.css"]
    code-fold: true
    code-overflow: wrap
    toc: true
    toc-expand: true
    toc-location: left
    lang: en-GB
    grid:
      sidebar-width: 250px
      body-width: 1200px
      margin-width: 250px
      gutter-width: 2em

execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true
  freeze: auto

knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 8.5

editor: source
---

```{r setup}
#| include: false
#| eval: true
#| cache: false
#| message: false

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

```{r global-variables}
#| include: false

source(here("00_model_functions.R"))

# Input paths
input_farm_coords_path <- here() %>% file.path("data", "_general_data", "farm_locations")
input_farm_sst_path <- here() %>% file.path("data", "_general_data", "SST")
input_species_param_path <- here() %>% file.path("data", "atlantic_salmon", "params")
input_feed_profile_path <- here() %>% file.path("data", "_general_data", "diets")

# Output paths
output_farm_data_path <- here() %>% file.path("outputs", "farm_data")
output_species_data_path <- here() %>% file.path("outputs", "species_data")
output_sens_data_path <- here() %>% file.path("outputs", "sensitivity_data")
output_farm_growth_data_path <- here() %>% file.path("outputs", "farm_growth_data")
output_cohort_growth_data_path <- here() %>% file.path("outputs", "cohort_growth_data")
data_analysis_path <- here() %>% file.path("outputs", "data_analysis")

# Filenames
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
sens_results_files <- file.path(output_sens_data_path) %>% str_subset("results")

species_params <- qread(species_params_file)
pop_params <- qread(pop_params_file)

feed_pal <- c("past" = brewer.pal(9, "Set1")[1], "reference" = brewer.pal(9, "Set1")[2], "future" = brewer.pal(9, "Set1")[3])
cohort_pal <- c("1" = brewer.pal(9, "Set1")[1], "2" = brewer.pal(9, "Set1")[2], "3" = brewer.pal(9, "Set1")[3], "all" = "black")
```

This document was last updated at `r format(Sys.time(), "%H:%M")` on `r format(Sys.time(), "%d %b, %Y")`.

# Methods
### Feeding rate

To detail here:

* Functional response to temperature (species-specific)
* Overfeeding parameters and how they come into play
* Deminstration of how feeding response and actual food given vary with temperature?

```{r fig-functional-response-to-temperature}
#| eval: false
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

df <- data.frame(water_temp = seq(-5, 26, 0.25))
df$feeding <- sapply(FUN = feeding_rate, X = df$water_temp, species_params = species_params)
df <- df %>% 
mutate(ing_pot = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * feeding,
       food_prov = NA)
for (i in 1:nrow(df)) {
  df$food_prov[i] <- food_prov_rate(pop_params = pop_params, rel_feeding = df$feeding[i], ing_pot = df$ing_pot[i], 
                                    ing_pot_10 = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * 0.1)
}
df$ing_pot <- df$ing_pot/maxna(df$ing_pot)
df$food_prov <- df$food_prov/maxna(df$ing_pot)

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

## Farm temperature forcings

```{r farm info}
#| code-summary: Get general farm info

farms_to_omit <- file.path(input_farm_coords_path, "atlantic_salmon_farms_to_omit.qs") %>% qs::qread()
farm_ts_data <- file.path(output_farm_data_path, "farm_ts_data.qs") %>% qs::qread()
farm_coords <- file.path(output_farm_data_path, "farm_coords.qs") %>% qs::qread() %>% 
    mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N") %>% as.factor())

farms_geometry <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread() %>% 
  rename(farm_ID = farm_id)

sorted_countries <- farm_coords %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  group_by(country) %>% 
  reframe(av_lat = mean(lat)) %>% 
  arrange(-av_lat) %>% 
  pull(country) %>% 
  unique()
```

```{r farm temperature data}
#| code-summary: Get the farm temperature forcings

farm_temp_means <- farm_ts_data %>% 
  group_by(farm_ID) %>% 
  reframe(mean_temp = mean(temp_c)) %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  dplyr::filter(!farm_ID %in% farms_to_omit)

overall_mean <- mean(farm_temp_means$mean_temp)
overall_sd <- sd(farm_temp_means$mean_temp)

rm(farm_ts_data)
```

Mean daily temperatures at the farms globally ranged from  from `r round(min(farm_temp_means$mean_temp), 1)`$^{\circ}$C to `r round(max(farm_temp_means$mean_temp), 1)`$^{\circ}$C, with an overall mean of `r round(overall_mean, 1)`$^{\circ}$C ($\pm$ `r round(overall_sd, 1)`$^{\circ}$C SD).
Farms with a mean temperature of $\leq$6$^\circ$C were excluded (`r as.integer(length(farms_to_omit))` farms). The total number of farms included in the analysis was therefore `r nrow(farm_temp_means)`.

```{r fig-farm-mean-temps}
#| fig-cap: |
#|   Potential options for showing farm site mean temperatures. Top left: Mean temperature by latitude across countries. Top right: Frequency distribution of mean temperatures by hemisphere. Bottom: global distribution of mean temperatures.
#| fig-height: 9
#| fig-width: 9.5

worldmap <- ne_countries(scale = "medium", returnclass = "sf")

p1 <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

p3 <- farm_temp_means %>% 
  mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N")) %>% 
  ggplot(aes(x = mean_temp, fill = hemisphere)) +
  geom_histogram(position = "identity", colour = "black", alpha = 0.5, binwidth = 0.5) +
  labs(x = expression("Mean temperature ("*degree*"C)"), y = "Frequency") +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

(p2 + p3) / p1 + plot_layout(heights = c(1, 1.5))
```

```{r patchwork map}
#| fig-cap: |
#|   Possibly use a patchwork map to show spatial results?
#| fig-height: 9
#| fig-width: 9.5

worldmap <- ne_countries(scale = "large", returnclass = "sf")

AUS_bounds <- c(144, 149.5, -39.75, -44)
CAN1_bounds <- c(-132, -122, 47.75, 54.25)
CAN2_bounds <- c(-70, -54, 43, 48.5)
CHI_bounds <- c(-77.5, -62.5, -56, -25)
EUR_bounds <- c(-26, 30, 51, 71)

p_ALL <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

no_margins <- function() {
  theme(legend.position = "none", 
        plot.margin = margin(0, 0, 0, 0),
        axis.title = element_blank())
}

p_Aus <- p_ALL + coord_sf(xlim = AUS_bounds[1:2], ylim = AUS_bounds[3:4], expand = FALSE) + no_margins()
p_Eur <- p_ALL + coord_sf(xlim = EUR_bounds[1:2], ylim = EUR_bounds[3:4], expand = FALSE) + no_margins()
p_Chi <- p_ALL + coord_sf(xlim = CHI_bounds[1:2], ylim = CHI_bounds[3:4], expand = FALSE) + no_margins()
p_Can <- p_ALL + coord_sf(xlim = CAN1_bounds[1:2], ylim = CAN1_bounds[3:4], expand = FALSE) + no_margins()
p_USA <- p_ALL + coord_sf(xlim = CAN2_bounds[1:2], ylim = CAN2_bounds[3:4], expand = FALSE) + no_margins()

subplot <- (p_Eur/p_Aus) + plot_layout(heights = c(1.5, 1))
finalplot <- (p_Can + p_USA) / (p_Chi + subplot) + 
  plot_layout(widths = c(1, 1.5), heights = c(1, 2)) &
  theme(plot.margin = margin(0, 0, 0, 0))
finalplot
```

## Production parameters

Salmon were grown for 548 days and were initialised Spring. This was on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` for farms in the southern hemisphere and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` for farms in the northern hemisphere. Cohorts were initialised anually, such that there was an overlap of 183 days where one cohort was nearing harvest and another was being initialised.

## Model parameterisation

```{r salmon params}
#| code-summary: Get parameter and sensitivity data
#| eval: true

params <- file.path(output_species_data_path, "sens_params.qs") %>% qs::qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

@tbl-params shows the parameter values used to model the growth of farmed Atlantic salmon.

| Parameter | Value | Units | Description | Reference | 
|:-------:|:----:|:----:|:----------------|:-------------------------|
| $\alpha$             | `r unname(params['alpha'])`      | -                        | Feeding catabolism coefficient | *Salvelinus alpinus*, *Oncorhynchus mykiss* (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | J g$^{-1}$               | Energy content of protein      | @baldan_r_2018 | 
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | J g$^{-1}$               | Energy content of lipid        | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | J g$^{-1}$               | Energy content of carbohydrate | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | J g O$_2^{-1}$           | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 |
| pk                   | `r unname(params['pk'])`         | d$^{-1}$                 | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | -                        | Fasting catabolism at 0 $^{^\circ}$C | @macnaughton_using_2019 |
| m                    | `r unname(params['m'])`          | -                        | Weight exponent for anabolism | Experimentally-derived alue over a range of *Oncorhynchus*, *Perca* and *Salmo* species (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| n                    | `r unname(params['n'])`          | -                        | Weight exponent for catabolism | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | -                        | Shape coefficient for the feeding function | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$\^{\circ}$C. Coefficient for mean daily energy intake of the moderately fast growing group, 0.4398 (mean fast group), 0.2214 (mean slow group) [@jonsson_thermal_2001] | 
| $T_{max}$            | `r unname(params['Tma'])`        | $^{\circ}$C              | Maximum lethal temperature | Juvenile *Oncorhynchus tshawytscha* (19.8 $\pm$ 0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) [@poletto_unusual_2017] |
| $T_{opt}$            | `r unname(params['Toa'])`        | $^{\circ}$C              | Optimal temperature | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{^\circ}$C [@jonsson_thermal_2001]. Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0 $\pm$ 14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$            | `r unname(params['Taa'])`        | $^{\circ}$C              | Lowest feeding temperature | Fish essentially stop feeding when below 2$^{\circ}$C [@vadboncoeur_lowering_2023] | 
| $\omega$             | `r unname(params['omega'])`      | g O$_2$ g$^{-1}$         | Oxygen consumption:weight loss ratio | Calibrated value from @baldan_r_2018 | 
| a                    | `r unname(params['a'])`          | J g tissue$^{-1}$        | Energy content of fish tissue | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| k                    | `r unname(params['k'])`          | -                        | Weight exponent for energy content | Alt value: 1.65175 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| eff                  | `r unname(params['eff'])`        | -                        | Food ingestion efficiency | Range: 95-97% ingestion efficiency [@uglem_does_2020] | 
| $\overline{W}$       | `r unname(params['meanW'])`      | g                        | Dry weight mean | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | g                        | Dry weight standard deviation | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\overline{F}$       | `r unname(params['overFmean'])`  | g g$^{-1}$ fish          | Overfeeding rate mean |    |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | g g$^{-1}$ fish          | Overfeeding rate standard deviation |    |
| mort                 | `r unname(params['mortmyt'])`    | d$^{-1}$                 | Natural mortality rate | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon {#tbl-params}

### Parameter sensitivity

Parameter sensitivity was tested by varying each parameter in @tbl-params by $\pm 10$% and calculating sensitivity as $S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}$, where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively. 
* All parameters in Table @tbl-params were tested across all farms (all temperatures)
* Used the full population of 5000 for the population parameters $\overline{W}$, $\Delta{W}$, $\overline{I_{max}}$, $\Delta{I_{max}}$, $\overline{F}$, and $\Delta{F}$, and a population of 100 in each farm for all other parameters.

```{r sensitivity data}
sens_weight <- output_sens_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("results") %>% 
  str_subset("weight") %>% 
  qs::qread()

sens_excr <- output_sens_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("results") %>% 
  str_subset("total_excr") %>% 
  qs::qread()

sens_data <- rbind(sens_weight, sens_excr) %>% 
  mutate(adj_param = factor(adj_param, levels = param_names$name))
```

Looking at @fig-param-sens-weight:

* The most sensitive parameter by far is $m$ (the weight exponent for anabolism), followed closely by $n$ the (weight exponent for catabolism)
* These two are also relatively uncertain, as $m$ was taken from *Salmo trutta* measurements in a fairly old paper [@broekhuizen_modelling_1994] and $n$ was taken from @baldan_r_2018 which did not specifically look at Atlantic salmon. There must be more up-to-date measurements for Atlantic salmon in particular.

```{r fig-param-sens-weight}
#| fig-cap: Impact of increasing each model parameter by 10% on final farm biomass (red) and total excreted nutrients (blue) over the whole production period.

sens_data %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens - sd_sens, ymax = mean_sens + sd_sens, fill = measure)) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black", alpha = 0.75) +
  # geom_errorbar(position = position_dodge(width = 0.95), width = 0.3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_y_continuous(breaks = seq(-30,30,5), limits = c(-10,30)) +
  scale_x_discrete(limits = rev(param_names$name), labels = rev(param_names$lab)) +
  scale_fill_manual(values = c(weight = "salmon", total_excr = "steelblue")) +
  # scale_fill_manual(values = c("weight" = "salmon", "uneat" = "steelblue", "excr" = "darkgreen")) +
  labs(x = "Adjusted parameter", y = "Sensitivity") +
  theme_classic() + 
  theme(legend.position = "none", 
        strip.text = element_blank(), 
        aspect.ratio = 0.8,
        text = element_text(size = 12),
        axis.title.y = element_blank()) +
  coord_flip()
```

## Modelled experimental feeds {#sec-feeds}

```{r feeds}
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>% qs::qread()

df <- feed_params[["reference"]]
contribP_1 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_1 <- sum(contribP_1)
digestP_1 <- sum(df[["Proteins"]]$digest * contribP_1)/tP_1

contribL_1 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_1 <- sum(contribL_1)
digestL <- sum(df[["Lipids"]]$digest * contribL_1)/tL_1

contribC_1 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_1 <- sum(contribC_1)
digestC_1 <- sum(df[["Carbohydrates"]]$digest * contribC_1)/tC_1

df <- feed_params[["past"]]
contribP_2 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_2 <- sum(contribP_2)
digestP_2 <- sum(df[["Proteins"]]$digest * contribP_2)/tP_2

contribL_2 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_2 <- sum(contribL_2)
digestL <- sum(df[["Lipids"]]$digest * contribL_2)/tL_2

contribC_2 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_2 <- sum(contribC_2)
digestC_2 <- sum(df[["Carbohydrates"]]$digest * contribC_2)/tC_2

df <- feed_params[["future"]]
contribP_3 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_3 <- sum(contribP_3)
digestP_3 <- sum(df[["Proteins"]]$digest * contribP_3)/tP_3

contribL_3 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_3 <- sum(contribL_3)
digestL <- sum(df[["Lipids"]]$digest * contribL_3)/tL_3

contribC_3 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_3 <- sum(contribC_3)
digestC_3 <- sum(df[["Carbohydrates"]]$digest * contribC_3)/tC_3
```

[Possibly replace this with a better summary of the ingredients that make up each feed, and report these stats in Results under Uneaten feed?]{.aside}

* Feeds came from [@cottrell_origins_nodate; @cottrell_global_2020].
* The reference feed is `r round(100*tP_1,1)`% protein, `r round(100*tL_1,1)`% lipid, `r round(100*tC_1,1)`% carbohydrate. 
* The past feed is `r round(100*tP_2,1)`% protein, `r round(100*tL_2,1)`% lipid, `r round(100*tC_2,1)`% carbohydrate. 
* The future feed is `r round(100*tP_3,1)`% protein, `r round(100*tL_3,1)`% lipid, `r round(100*tC_3,1)`% carbohydrate. 
* Within the reference feed, protein is overall `r round(100*digestP_1,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_1,1)`% digestible.
* Within the past feed, protein is overall `r round(100*digestP_2,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_2,1)`% digestible.
* Within the future feed, protein is overall `r round(100*digestP_3,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_3,1)`% digestible.

## Model population and biomass

```{r fig-cohorts}
#| fig-cap: | 
#|   Example of how the overlapping cohorts combine into a single farm production curve (in this case biomass). The left plot shows the farm biomass produced by each cohort from initialisation to harvest, while the right plot shows the combined farm biomass over the course of a single production period of 548 days.
#| eval: true

fnms <- output_farm_growth_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("reference")

df <- qs::qread(fnms[1])[["biomass_stat"]] %>% 
  as.data.frame() %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t")) %>% 
  select(-farm_ID)

df <- rbind(
  df %>% mutate(cohort = 1),
  df %>% mutate(t = t+365, cohort = 2),
  df %>% mutate(t = t+730, cohort = 3)
) %>% 
  mutate(t = (t-min(t))+1,
         split = "Split by cohort")

min_t <- min(df$t[df$cohort == 2])
max_t <- max(df$t[df$cohort == 2])

df_2 <- df %>% 
  filter(t %in% min_t:max_t) %>% 
  mutate(sd = sd/mean) %>% 
  group_by(t) %>% 
  reframe(sd = sqrt(sum(sd^2)),
          mean = sum(mean),
          sd = sd*mean) %>% 
  mutate(cohort = "all",
         split = "Totalled")

df <- rbind(df, df_2) %>% 
  mutate(cohort = as.factor(cohort))

ggplot(df, aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  facet_grid(cols = vars(split)) +
  theme_classic() +
  scale_x_continuous(breaks = seq(0, 1600, 120)) +
  scale_colour_manual(values = cohort_pal) +
  theme(legend.position = "none") +
  labs(y = "Farm biomass", x = "Day of production")
```

```{r pop-biom-data}
coho_biom <- file.path(data_analysis_path, "biomass_produced_comparison.qs") %>% 
  qs::qread()
harv_size <- file.path(output_farm_data_path, "farm_harvest_size.qs") %>% 
  qread()

coho_biom_summ <- coho_biom %>% 
  mutate(perc_diff = 100*(mean - tonnes_per_farm)/tonnes_per_farm)
```

Each farm's final population was calculated by dividing the production tonnage required by the average modelled individual harvest size for each farm (when feed reference feed). A population timeseries was the generated using the standard mortality rate (`r unname(params['mortmyt'])` d$^{-1}$). Using this method, a farm's 'population' is not a good representation of the actual number of individuals being farmed, but it ensures each farm location is producing the correct tonnage while allowing comparisons between the different feeds.

```{r fig-harvest-size, class.chunk = "aside"}
#| fig-cap: |
#|   Frequency distribution of harvest sizes used to create the population timeseries for each farm, as predicted by the fish-growth model.

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  theme_classic() +
  theme(
    legend.position = "none", 
    aspect.ratio = 0.75
  )
```

By this method, farm biomass as predicted by the model slightly overestimated farm tonnage. Predicted farm biomass was `r round(min(coho_biom_summ$perc_diff),1)`-`r round(max(coho_biom_summ$perc_diff),1)` higher than tonnage produced (@fig-biomass-produced). This variation was not correlated with mean farm temperature or total tonnage produced (@fig-biomass-variation).

```{r fig-biomass-variation, class.chunk = "aside"}
plot(coho_biom_summ$temp_c, coho_biom_summ$perc_diff)
plot(coho_biom_summ$tonnes_per_farm, coho_biom_summ$perc_diff)
```

```{r fig-biomass-produced, class.chunk = "aside"}
#| fig-cap: Modelled vs observed salmon biomass produced

coho_biom %>% 
  ggplot(aes(x = tonnes_per_farm, y = mean, colour = country)) +
  geom_point(size = 2.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_classic() +
  scale_x_continuous(limits = c(0,1700)) +
  scale_y_continuous(limits = c(0,1700)) +
  scale_colour_brewer(palette = "Set1") +
  labs(y = "Modelled biomass produced (t)", x = "Observed biomass produced (t)") +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())
```

## Impacts of nutrient input

*To detail here:*

* How to convert feed and uneaten faeces into nutrients - just nitrogen? Carbon too? If just nitrogen, then can probably use the standard N = 6.25 x protein conversion to compare the feeds (lipids and carbohydrates have negligible protein). If doing carbon too, will need to find conversion rates for all of them, possibly splitting carbs/lipids into complex/simple, saturated/unsaturated, etc.
* Source of species vulnerability layers

# Results

In general, data outputs comes in the following dimensions:

* Location -> temperature -> farm_ID
* Uneaten feed + excreted faeces = total inputs (g)
* Day of production (impacted by location)

* Main point is the difference between feeds (kg/t salmon).
* Also interesting to see if difference between feeds varies geographically (correlates with latitude, or maybe mean/median/max temperature?)

```{r weight and biomass data}
farmrun_files <- file.path(output_cohort_growth_data_path) %>% 
  list.files(full.names = T)

qs::qread(farmrun_files[1]) %>% names()
weight <- purrr::map(farmrun_files, function(f) {qs::qread(f)[["weight_stat"]]}) %>% bind_rows()
biomass <- purrr::map(farmrun_files, function(f) {qs::qread(f)[["biomass_stat"]]}) %>% bind_rows()
```

## Uneaten feed

```{r uneaten feed data}
#| code-summary: Load and prep uneaten feed data

uneaten_total <- farmrun_files %>% 
  purrr::map(function(f) {qs::qread(f)[["total_uneat_stat"]]}) %>% 
  bind_rows() %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(uneat_mean = mean.x, biomass_mean = mean.y,
         uneat_sd = sd.x, biomass_sd = sd.y) %>% 
  mutate(uneat_sd = uneat_sd/uneat_mean,
         biomass_sd = biomass_sd/biomass_mean,
         uneat_biom_mean = uneat_mean/biomass_mean,
         uneat_biom_sd = sqrt(uneat_sd^2+biomass_sd^2)*uneat_biom_mean)

uneaten_total_stats <- uneaten_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(uneat_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         max = max %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         mean = mean %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))
```

### Amount

* Total reference uneaten feed across a cohort (548 days) ranged from `r uneaten_total_stats$min[uneaten_total_stats$feed == "reference"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "reference"]` g kg$^{-1}$, with a mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "reference"]`.
* There was very little difference between the past and future feeds and the reference feed. Uneaten past feed ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "past"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "past"]` g kg$^{-1}$ (mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "past"]`) while uneaten future feed ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "future"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "future"]` g kg$^{-1}$ (mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "future"]`).
* @fig-total-uneaten-perday shows the range of total uneaten feed per day across countries.

```{r fig-total-uneaten-perday}
#| fig-width: 9.5

uneaten_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(uneat_biom_mean = uneat_biom_mean %>% set_units("g g_fish-1 d-1") %>% set_units("g kg_fish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = uneat_biom_mean, fill = feed)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Uneaten feed", x = "Country")
```

```{r uneaten composition}
#| eval: false

uneaten_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_uneat_stat"]]}) %>% bind_rows()
uneaten_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_uneat_stat"]]}) %>% bind_rows()
uneaten_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_uneat_stat"]]}) %>% bind_rows()

uneaten_all <- uneaten_total %>% 
  select(-contains("biom")) %>% 
  merge(uneaten_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd")))
```

* The composition of the uneaten feed is (obviously) the same as the total feed composition (@sec-feeds).

## Excreted faeces

```{r excreted faeces data}
#| code-summary: Load and prep all excretion data

# The following reads in all the total excretion data and converts the mean and sd values into values/biomass
excreted_total <- farmrun_files %>% 
  purrr::map(function(f) {qs::qread(f)[["total_excr_stat"]]}) %>% 
  bind_rows() %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(excr_mean = mean.x, biomass_mean = mean.y,
         excr_sd = sd.x, biomass_sd = sd.y) %>% 
  mutate(excr_sd = excr_sd/excr_mean,
         biomass_sd = biomass_sd/biomass_mean,
         excr_biom_mean = excr_mean/biomass_mean,
         excr_biom_sd = sqrt(excr_sd^2+biomass_sd^2)*excr_biom_mean)

excreted_total_stats <- excreted_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(excr_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         max = max %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1"),
         mean = mean %>% set_units("g g_fish-1") %>% set_units("g kg_fish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))
```

* Total reference excreted faeces across a cohort (548 days) ranged from `r excreted_total_stats$min[excreted_total_stats$feed == "reference"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "reference"]` g kg$^{-1}$, with a mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "reference"]`.
* There was very little difference between the past and future feeds and the reference feed. Uneaten past feed ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "past"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "past"]` g kg$^{-1}$ (mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "past"]` g kg$^{-1}$) while uneaten future feed ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "future"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "future"]` g kg$^{-1}$ (mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "future"]` g kg$^{-1}$).
* @fig-total-excreted-perday shows the range of total excreted faeces per day across countries.

```{r fig-total-excreted-perday}
#| fig-width: 9

excreted_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(excr_biom_mean = excr_biom_mean %>% set_units("g g_fish-1 d-1") %>% set_units("g kg_fish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = excr_biom_mean, fill = feed)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Excreted faeces", x = "Country")
```

### Amount

```{r fig-total-excr-diff_perc, class.chunk = "aside"}
#| fig-cap: Mean % difference in excreted faeces of the two experimental feeds (compared to the reference feed) with increasing overall excretion.

df <- excreted_total %>% 
  select(-contains(c("sd", "biom"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kg_fish-1 d-1"),
         diff_perc = drop_units(diff/reference)) %>% 
    merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t - t_start + 1)

summ <- df %>%
  group_by(hemisphere, feed) %>% 
  reframe(mean = 100*meanna(diff_perc) %>% round(3),
          min = 100*meanna(diff_perc) %>% round(3),
          max = 100*maxna(diff_perc) %>% round(3))

df %>% 
  ggplot(aes(x = reference, y = diff_perc, colour = hemisphere)) +
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(~feed) +
  theme_classic()
```

The difference in total excretion for the past feed ranged from 
`r summ$min[summ$feed == "past"]`-`r summ$max[summ$feed == "past"]`% across all farms globally, and for the future feed ranged from 
`r summ$min[summ$feed == "future"]`-`r summ$max[summ$feed == "future"]`%.

```{r fig-total-excr-diff_perc_2, class.chunk = "aside"}
#| fig-cap: Mean % difference in excreted faeces of the two experimental feeds (compared to the reference feed) over time in one production cycle.

df %>% 
  group_by(prod_day, feed, hemisphere) %>% 
  reframe(diff_perc_mean = meanna(diff_perc),
          diff_perc_sd = sdna(diff_perc)) %>% 
  ggplot(aes(x = prod_day, y = diff_perc_mean, ymin = diff_perc_mean-diff_perc_sd, ymax = diff_perc_mean+diff_perc_sd, 
             colour = as.factor(hemisphere), fill = as.factor(hemisphere))) +
  geom_line() + 
  geom_ribbon(alpha = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(~feed) +
  theme_classic()
```

However, the actual difference in excreted faeces varied by feed, time in the production cycle, and location of the farm [@fig-total-excr-act_diff].

```{r fig-total-excr-act_diff}
#| fig-width: 6.25
#| fig-height: 5.25
#| fig-cap: Mean actual difference in excreted faeces of the two experimental feeds (compared to the reference feed) over time in one production cycle.

df <- excreted_total %>% 
  select(-contains(c("sd", "excr_mean", "biomass"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_biom_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kg_fish-1 d-1"),
         diff = set_units(diff, "kg t_fish-1 d-1")) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t-t_start+1)

summ <- df %>%
  group_by(hemisphere, feed) %>% 
  reframe(mean = meanna(diff) %>% round(3),
          min = meanna(diff) %>% round(3),
          max = maxna(diff) %>% round(3))

df %>% 
  group_by(hemisphere, feed, prod_day) %>% 
  reframe(mean_diff = meanna(diff),
          sd_diff = sdna(diff) %>% set_units("kg t_fish-1 d-1")) %>% 
  ggplot(aes(x = prod_day, y = mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff, colour = hemisphere, fill = hemisphere)) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(~feed) +
  theme_classic()
```

### Composition

```{r excreted composition}
#| eval: true

excreted_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_excr_stat"]]}) %>% bind_rows()
excreted_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_excr_stat"]]}) %>% bind_rows()
excreted_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_excr_stat"]]}) %>% bind_rows()

excreted_all <- excreted_total %>% 
  select(-contains("biom")) %>% 
  merge(excreted_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd"))) %>% 
  rename(protein = mean.x, lipids = mean.y, carbohydrates = mean) %>% 
  mutate(protein = protein/excr_mean %>% set_units("d-1"),
         lipids = lipids/excr_mean %>% set_units("d-1"),
         carbohydrates = carbohydrates/excr_mean %>% set_units("d-1")) %>% 
  select(-excr_mean)
```

Unlike the uneaten feed, the composition of excreted faeces differs from that of the incoming feed.

* Composition does not change through time or with increasing biomass (because digestibility does not change with fish size or temeprature)

```{r composition-throughtime}
#| eval: false

excreted_all %>% 
  pivot_longer(names_to = "macro", values_to = "perc", cols = c("protein", "lipids", "carbohydrates")) %>% 
  mutate(macro = factor(macro, levels = c("protein", "lipids", "carbohydrates"))) %>% 
  group_by(farm_ID, feed, macro) %>% 
  reframe(sd = sd(perc) %>% set_units("d-1"),
          mean = mean(perc)) %>% 
  ggplot(aes(x = farm_ID, colour = feed, fill = feed, y = 100*mean, ymin = 100*(mean-sd), ymax = 100*(mean+sd))) +
  geom_point(size = 2) +
  geom_errorbar(width = 0.5) +
  scale_color_manual(values = feed_pal) +
  scale_fill_manual(values = feed_pal) +
  facet_grid(rows = vars(macro)) +
  theme_classic() +
  labs(x = "Macronutrient", y = "Mean composition of faeces")
```

| Macro | Reference | Past | Future |
|:-----------:|:---:|:---:|:---:|
| Protein | `r tP_1` |  `r tP_2` |  `r tP_3` | 
| Lipids | `r tL_1` | `r tL_2` |  `r tL_3` | 
| Carbohydrates | `r tC_1` | `r tL_2` |  `r tL_3` |

## Total nutrients

# Statistical analysis & code availability

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `terra` [@terra], `targets` [@targets], `future` [@future], `furrr` [@furrr], `renv` [@renv], and the `tidyverse` [@tidyverse].
For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].

# References