### 2025-06-13 02_extracting_production_locations.qmd

---
title: "Extracting production locations for each species to obtain environmental forcings"
format: html
editor: source
---

## Set up

```{r}

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(qs)
library(countrycode)
library(terra)

source("src/dir.R")

species <- c("atlantic_salmon", 
             "general_salmonid", 
             "european_seabass", 
             "gilthead_seabream", 
             "large_yellow_croaker", 
             "japanese_seabass",
             "japanese_amberjack",
             "general_grouper",
             "pompano",
             "milkfish")

```

Isolate production data for species of interest

```{r}

aquaculture_raw <- read_csv(file.path(rdsi_dir, "raw_data/fao/FAO_fishstat_2020/global_aquaculture_production_1950_2020.csv"))

(aquaculture_prod <- 
  aquaculture_raw |> 
  clean_names() |> 
  select(c(1:7, starts_with("x"))) |> 
  pivot_longer(names_to = "year", values_to = "production", cols= -c(1:7)) |> 
  mutate(year = gsub(pattern="x", replacement = "", year)) |> 
  filter(year %in% c(2015:2020)) |> 
  group_by(asfis_species_name_2, asfis_species_name_3, environment_name, year) |> 
  summarise(production = sum(production)) |> 
    ungroup() |> 
     filter(environment_name=="Marine" & 
           !asfis_species_name_3 %in% c("Clams, cockles, arkshells", 
                                        "Abalones, winkles, conchs",
                                        "Mussels", 
                                        "Oysters",
                                        "Shrimps, prawns",
                                        "Crabs, sea-spiders",
                                         "Freshwater molluscs",
                                        "Squids, cuttlefishes, octopuses",
                                        "Pearls, mother-of-pearl, shells",
                                        "Miscellaneous aquatic plants",
                                        "Scallops, pectens"),
         !grepl("seaweeds" , asfis_species_name_3),
         !grepl("molluscs|cucumber|invertebrates", asfis_species_name_2)) |> 
    filter(!asfis_species_name_2 %in% c("Marine fishes nei")) |> 
     group_by(asfis_species_name_2) |> 
     summarise(production = mean(production)) |> 
  arrange(-production) |> 
    mutate(cum_prop = cumsum(production)/sum(production),
           prop = production/sum(production))
)

spp_names <- 
  aquaculture_prod |> 
  filter(asfis_species_name_2 %in% c("Atlantic salmon", 
                                     "Rainbow trout", 
                                     "European seabass", 
                                     "Large yellow croaker",
                                     "Gilthead seabream",
                                     "Coho(=Silver) salmon",
                                     "Japanese seabass", 
                                     "Japanese amberjack", 
                                     "Amberjacks nei",
                                     "Pompano", 
                                     "Milkfish") | 
           grepl(pattern = "rouper", asfis_species_name_2) & production > 1) |> 
  pull(asfis_species_name_2)

(mariculture_spp_prod <- 
    aquaculture_raw |> 
    clean_names() |> 
    select(c(1:7, starts_with("x"))) |> 
    pivot_longer(names_to = "year", values_to = "production", cols= -c(1:7)) |> 
    mutate(year = gsub(pattern="x", replacement = "", year)) |> 
    filter(year %in% c(2017)) |> 
    filter(!grepl("Inland waters", fao_major_fishing_area_name)  & 
             !asfis_species_name_3 %in% c("Clams, cockles, arkshells", 
                                          "Abalones, winkles, conchs",
                                          "Mussels", 
                                          "Oysters",
                                          "Shrimps, prawns",
                                          "Crabs, sea-spiders",
                                          "Freshwater molluscs",
                                          "Squids, cuttlefishes, octopuses",
                                          "Pearls, mother-of-pearl, shells",
                                          "Miscellaneous aquatic plants",
                                          "Scallops, pectens"),
           !grepl("seaweeds" , asfis_species_name_3),
           !grepl("molluscs|cucumber|invertebrates", asfis_species_name_2)) |> 
    group_by(country_name, asfis_species_name_2, asfis_species_name_3, fao_major_fishing_area_name, fao_major_fishing_area_code, year) |> 
    summarise(production = sum(production)) |> 
    ungroup() |> 
    mutate(asfis_species_name_2 = case_when(asfis_species_name_2 == "Marine fishes nei" & country_name == "Viet Nam" ~ "Groupers nei",
                                            TRUE ~ asfis_species_name_2)) |> 
    mutate(iso3c = countrycode(sourcevar = country_name, origin = "country.name", destination = "iso3c", warn =TRUE)) |> 
    group_by(iso3c, country_name, year) |> 
    nest() |> 
    mutate(national_production = map(data, ~(sum(.$production)))) |> 
    unnest(cols = c(data, national_production)) |> 
    ungroup() |> 
    filter(asfis_species_name_2 %in% spp_names) |> 
    mutate(prop_production = production/national_production) |> 
    filter(!asfis_species_name_2 %in% c("Marine fishes nei")) |> 
  mutate(model_name = case_when(grepl("rouper", asfis_species_name_2) ~ "general_grouper", 
                                asfis_species_name_2 %in% c("Coho(=Silver) salmon", "Rainbow trout") ~ "general_salmonid",
                                asfis_species_name_2 %in% c("Amberjacks nei") ~ "japanese_amberjack",
                                TRUE ~ gsub(pattern = " ", replacement = "_", tolower(asfis_species_name_2)))) |> 
  relocate(model_name, .before = asfis_species_name_2) |> 
  filter(prop_production>0) |> 
    mutate(farm_name = case_when(model_name %in% c("atlantic_salmon", "general_salmonid") ~ "Salmonidae fish",
                                 TRUE ~ "General marine fish"))
    )

```

Pull in farm locations, add FAO Major fishing regions

```{r}
farms_sf <- qread(file = "data/_general_data/farm_locations/locations.qs") |> 
  st_as_sf(coords = c("X", "Y"), crs = "EPSG:4326") |> 
  mutate(row_num = row_number()) |> 
  group_by(row_num) |> 
  group_split()

FAO_shp <- read_sf(file.path(rdsi_dir, "raw_data/fao/FAO_AREAS_CWP_NOCOASTLINE")) |> filter(F_LEVEL == "MAJOR") |> select(F_CODE)

#test function
#this_farm <- farms_sf[[1]]

# define cores for parallel process

cores <- parallel::detectCores()-2

#  process set up

future::plan(strategy = "multisession", workers = cores)

#intersect farm points with FAO fishing code polygons

farm_fao_intersect <-

  furrr::future_map(.x = farms_sf, .f = \(this_farm){ # initiate parallel map function

  this_farm_fao_code <- st_intersection(this_farm, FAO_shp)

  return(this_farm_fao_code)

}) |> bind_rows() #combine to data frame

qsave(x = farm_fao_intersect, file = here("data/_general_data/farm_locations/locations_w_fishing_areas.qs"))

```

Check that farm locations for each country sit in the correct FAO fishing area - flag those where there is a misalignment

```{r}

# filter country level production for all species in non-inland waters where production is greater than 0 tonnes.

mariculture_prod_all <- 
    aquaculture_raw |> 
    clean_names() |> 
    select(c(1:7, starts_with("x"))) |> 
    pivot_longer(names_to = "year", values_to = "production", cols= -c(1:7)) |> 
    mutate(year = gsub(pattern="x", replacement = "", year)) |> 
    filter(year %in% c(2017)) |> 
    filter(!grepl("Inland waters", fao_major_fishing_area_name)  & 
             !asfis_species_name_3 %in% c("Clams, cockles, arkshells", 
                                          "Abalones, winkles, conchs",
                                          "Mussels", 
                                          "Oysters",
                                          "Shrimps, prawns",
                                          "Crabs, sea-spiders",
                                          "Freshwater molluscs",
                                          "Squids, cuttlefishes, octopuses",
                                          "Pearls, mother-of-pearl, shells",
                                          "Miscellaneous aquatic plants",
                                          "Scallops, pectens"),
           !grepl("seaweeds" , asfis_species_name_3),
           !grepl("molluscs|cucumber|invertebrates", asfis_species_name_2)) |> 
    group_by(country_name, asfis_species_name_2, asfis_species_name_3, fao_major_fishing_area_name, fao_major_fishing_area_code, year) |> 
    summarise(production = sum(production)) |> 
    ungroup() |> 
    mutate(iso3c = countrycode(sourcevar = country_name, origin = "country.name", destination = "iso3c", warn =TRUE)) |> 
    group_by(iso3c, country_name, year) |> 
    nest() |> 
    mutate(national_production = map(data, ~(sum(.$production)))) |> 
    unnest(cols = c(data, national_production)) |> 
    ungroup() |> 
  filter(production>0) |> 
  filter(!asfis_species_name_2 %in% c("Marine fishes nei")) |> 
   mutate(model_name = case_when(grepl("rouper", asfis_species_name_2) ~ "general_grouper", 
                                asfis_species_name_2 %in% c("Coho(=Silver) salmon", "Rainbow trout") ~ "general_salmonid",
                                TRUE ~ gsub(pattern = " ", replacement = "_", tolower(asfis_species_name_2)))) |> 
  relocate(model_name, .before = asfis_species_name_2) |> 
  mutate(farm_name = case_when(model_name %in% c("atlantic_salmon", "general_salmonid") ~ "Salmonidae fish",
                                 TRUE ~ "General marine fish")) |> 
 filter(asfis_species_name_2 %in% spp_names) |> 
  mutate(prop_production = production/national_production)

#Check that farm locations in mapped data are in the same FAO fishing regions as the production data

farm_fao_intersect <- qread(here("data/_general_data/farm_locations/locations_w_fishing_areas.qs"))

farm_list <- 
  farm_fao_intersect |> 
  group_by(iso3c, country, species_group) |> 
  group_split()

#Test function

this_country_region <- farm_list[[4]]

(fao_area_code_check <- 
  map_df(farm_list, \(this_country_region){

    farm_country_name <- this_country_region$country |> unique()
    farm_iso3c <- this_country_region$iso3c |> unique()
    farm_fao_code <- this_country_region$F_CODE |> unique() |> sort()
    farm_spp_grp <- this_country_region$species_group |> unique()

    prod_data <- 
      mariculture_prod_all |> 
      filter(iso3c == farm_iso3c & 
               fao_major_fishing_area_code %in% farm_fao_code & 
               country_name == farm_country_name & 
               farm_name == farm_spp_grp)

    prod_spp_group <- prod_data$farm_name |> unique()

    prod_fao_code <- prod_data$fao_major_fishing_area_code |> unique() |> sort()

    return(
      tibble(iso3c = farm_iso3c, 
             countryname = farm_country_name, 
             farm_spp_group = farm_spp_grp, 
             mapped_fishing_areas = paste(farm_fao_code, collapse = ' '), 
             production_fishing_areas = paste(prod_fao_code, collapse = ' ')) |> 
             mutate(same_fishing_area = mapped_fishing_areas == production_fishing_areas))

}) |> filter(!production_fishing_areas == "")
)

```

Assign species to individual farms

```{r}
farm_list <- 
  farm_fao_intersect |> 
  group_by(iso3c, country, species_group, F_CODE) |> 
  group_split()

map_spp_to_locations_list <-

  map(.x = farm_list, .f = \(this_country_region){

  this_iso3c <- this_country_region$iso3c |> unique()
  this_fao_code <- this_country_region$F_CODE |> unique()
  this_species_group <- this_country_region$species_group |> unique()

  message("Processing locations for ", this_species_group, " in ", this_iso3c, " and FAO fishing area ", this_fao_code)

  this_n_farms <- this_country_region |> nrow()

  this_country_region_spp_production_list <- 
    mariculture_spp_prod |> 
    filter(iso3c == this_iso3c, farm_name == this_species_group, fao_major_fishing_area_code == this_fao_code) |> 
    group_by(model_name, national_production) |> 
    summarise(production = sum(production),
              prop_production = production/national_production) |> 
    ungroup() |> 
    distinct() |> 
    group_by(model_name) |> 
    group_split()

  farm_locations_w_model_spp <- 
    map_df(.x = this_country_region_spp_production_list, .f = \(this_species_prop){

    this_spp_n_farms <- round(this_n_farms* this_species_prop$prop_production)

    set.seed(1234)
    indices <- sample(nrow(this_country_region), size = this_spp_n_farms)

    this_spp_locations <- this_country_region[indices,]
    this_spp_locations$model_name <- this_species_prop$model_name

    this_country_region <- this_country_region[-indices,]

     return(this_spp_locations)

  })

  return(farm_locations_w_model_spp)

})

map_spp_to_locations <- bind_rows(lapply(map_spp_to_locations_list, FUN = function(element) if(nrow(element) == 0) NULL else element)) #removes the NULL record returns

qsave(x = map_spp_to_locations, file = here("data/_general_data/farm_locations/locations_w_species_fao_area.qs"))
```

Stocking densities from the tonnage per farm and harvest size.

```{r}

farms_w_species <- qread("data/_general_data/farm_locations/locations_w_species_fao_area.qs")

harvest_sizes <- read_csv("data/_general_data/harvest_sizes/all_harvest_sizes.csv") |> select(-harvest_size_g)

production_cycle <- read_csv("data/_general_data/production_cycles/production_cycle.csv") |> select(species, days)

mortality_rates <- read_csv("data/_general_data/mortality_rates/all_mortality_rates.csv")

farms_w_stocking <- 
  farms_w_species |> 
  left_join(harvest_sizes, by = c("model_name" = "species")) |> 
  left_join(production_cycle, by = c("model_name" = "species")) |> 
  left_join(mortality_rates, by = c("model_name" = "species")) |> 
  mutate(harvest_n = tonnes_per_farm/harvest_size_t) |> 
  mutate(stocking_n = harvest_n * exp(daily_mort_rate*days))

qsave(x = farms_w_stocking, file =  "data/_general_data/farm_locations/locations_w_species_fao_area_stocking.qs")

```