### 2025-06-18 07_species_layers.qmd

---
title: "07_species_layers"
format: html
editor: source
---

```{r}
library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
library(here)
library(terra)
library(sf)
library(qs)
library(ggplot2)
library(countrycode)
library(terra)
library(tidyterra)
library(cowplot)
library(furrr)
library(rnaturalearth)
library(aquamapsdata) # https://raquamaps.github.io/aquamapsdata/articles/intro.html
library(worrms)
library(grid)

source("00_dirs.R")

overwrite <- F
```

# Base rasters for resolution

Create a blank raster with enough resolution to get farms in there (100m)

```{r base-raster}
base_rast <- rast(res = 0.1/111.32, extent = ext(-180, 180, -90, 90), crs = "EPSG:4326")

inset_boxes <- list(
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 72),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -44, -39.75)
)

inset_rasters <- list(
  CAN1 = crop(base_rast, ext(inset_boxes[["CAN1"]])),
  CAN2 = crop(base_rast, ext(inset_boxes[["CAN2"]])),
  EUR  = crop(base_rast, ext(inset_boxes[["EUR"]])),
  CHI  = crop(base_rast, ext(inset_boxes[["CHI"]])),
  AUS  = crop(base_rast, ext(inset_boxes[["AUS"]]))
)
```

# Farm circles

Give each farm a basic "circle of influence" - currently set at 2km. Possibility to reduce this/create a dispersion pattern later.

```{r farm-circles}
farm_circles_file <- file.path(output_farm_data_path, "farm_circles_2km.tif")

farm_circles_rast <- if (!file.exists(farm_circles_file) | overwrite) {
  # Read in farm locations (sf objects)
  farm_locs <- file.path(output_farm_data_path, "farm_geometry.qs") %>% 
    qread()

  # Create circular buffers for crude (~1km) area of impact and merge overlapping areas
  farm_circles <- st_sf(farm_id = farm_locs$farm_id, geometry = farm_locs$geometry) %>%
    st_transform(crs = crs(base_rast)) %>%  # transform to match background N projection
    st_buffer(dist = 2000) %>%  # 2 km radius circles - for maximum possible impact (temporary for testing method)
    st_union() %>%  # merge overlapping polygons
    st_sf() %>%  # convert back to sf object
    mutate(id = 1)  # add single ID for rasterization

  farm_rast <- rasterize(vect(farm_circles), base_rast, field = "id") # this takes a long time because the base raster is so fine
  writeRaster(farm_rast, farm_circles_file, overwrite = T)
  farm_rast
} else {
  rast(farm_circles_file)
}
```

Also save the individual areas of farm clusters so later species-overlaying is a bit faster.

```{r small-bb-farm-circles}
fa_list <- file.path(output_farm_data_path, sprintf("farm_circles_2km_inset_%s.tif", 1:5))
for (i in 1:5) {
  if (!file.exists(fa_list[i]) | overwrite) {
    fa_rast <- crop(farm_circles_rast, inset_rasters[[i]])
    writeRaster(fa_rast, fa_list[i])
  }
}

# Testing out different resolutions for the farm_circles
fa_rast_100m <- rast(fa_list[4])
fa_rast_250m <- aggregate(fa_rast_100m, fact = 2.5, fun = "modal")
fa_rast_500m <- aggregate(fa_rast_100m, fact = 5, fun = "modal")
fa_rast_1000m <- aggregate(fa_rast_100m, fact = 10, fun = "modal")

fa_ggplot <- function(df) {
  ggplot() +
    geom_spatraster(data = df, mapping = aes(fill = id), na.rm = T) +
    scale_fill_viridis_c(na.value = "transparent") +
    coord_sf(expand = FALSE) +
    theme_void()
}

plot_grid(
  fa_ggplot(fa_rast_100m),
  fa_ggplot(fa_rast_250m),
  fa_ggplot(fa_rast_500m),
  fa_ggplot(fa_rast_1000m),
  ncol = 2
)
```

# Species vulnerabilities

Using trait-based vulnerability data downloaded from @ohara_code_2021 [@butt_trait-based_2022].

``` {r vulnerability data}
vulnerabilities <- file.path(input_spec_layers_path, "spp_gp_vuln_w_distribution.csv") %>% 
  read.csv() %>% 
  mutate(spp_gp = str_to_sentence(spp_gp)) %>% 
  filter(stressor == "eutrophication_nutrient_pollution") %>% 
  mutate(taxon = as.factor(taxon)) %>% 
  mutate(spp_gp = str_remove(spp_gp, "\\s*\\([^)]*\\)")) %>%  # remove non-accepted names in parentheses
  filter(exposure_mod != 0) %>% 
  select(-contains("sd"), -stressor)

ggplot(vulnerabilities, aes(x = taxon, y = vuln, fill = taxon)) +
  geom_boxplot() +
  coord_flip()
```

## Species ranges

```{r aquamaps database}
# download_db(force = FALSE) # This step downloads the Aquamaps database, ~ 2GB. Only need to do this once
default_db("sqlite") # Need this step in order to set up to query database
```

```{r aquamaps-species-keys}
vuln_am_keys <- split(vulnerabilities, vulnerabilities$spp_gp) %>% 
  lapply(function(vdf) {
    keys <- am_search_fuzzy(vdf$spp_gp) %>% 
      mutate(key = as.character(key),
             terms = as.character(terms),
             spp_gp = vdf$spp_gp,
             genus_species = str_extract(terms, "^\\S+\\s+\\S+"))
    merge(vdf, keys, by = "spp_gp")
  }) %>% 
  bind_rows() %>% 
  # The vulnerabilities are applied at different levels depending on data available, which can affect distribution lookups
  mutate(
    match_level = case_when(
      genus_species == spp_gp ~ "species",
      str_detect(str_to_lower(genus_species), str_to_lower(spp_gp)) ~ "genus",
      str_detect(spp_gp, "ae$") ~ "family",
      str_detect(str_to_lower(terms), str_to_lower(spp_gp)) & !str_detect(str_to_lower(genus_species), str_to_lower(spp_gp)) ~ "common",
      str_detect(str_to_lower(terms), str_to_lower(spp_gp)) ~ "other_taxon",
      T ~ NA
    ),
    match_level = factor(match_level, levels = c("species", "genus", "family", "other_taxon", "common"))
  ) %>% 
  arrange(match_level) %>% 
  # Keep only one key (aquamaps dataset) per species, but there may be multiple vulnerabilities
  distinct(key, .keep_all = T)
```

### Get species rasters

```{r species-rasters}
all_keys <- vuln_am_keys$key
for (i in seq_along(all_keys)) {
  ras <- am_raster(key = all_keys[i]) %>% rast()
  names(ras) <- "probability"
  writeRaster(ras, file.path(input_aquamaps_path, str_c(all_keys[i], ".tif")), overwrite = T)
}
```

Convert to Gall projection for equal area?

# Calculate overlapping area

```{r species-area-setup}
vuln_am_keys <- vuln_am_keys %>% 
  mutate(total_species_area = NA,
         farm_overlap_area = NA,
         file_nm = file.path(input_aquamaps_path, str_c(key, ".tif"))) %>% 
  select(-terms)

fa_rasts <- list.files(output_farm_data_path, full.names = T, pattern = "farm_circles_2km_inset") %>% 
  as.list() %>% 
  lapply(function(fnm){
    rast(fnm) %>% aggregate(fact = 10, fun = "modal") # use 1km resolution for speed
  })
fa_exts <- lapply(fa_rasts, ext) %>% unlist()
```

Get the total overlap between the farm buffer circles (1km radius) and the species range (presence/absence). The data from aquamaps gives a probability of species presence (0-1) but that is not incorporated here (yet).

I tried to do this in parallel but `terra` has issues with parallel processing - couldn't get it to work. But it only takes ~25 minutes (with all the speed measures taken above).

```{r species-area}
for (sp in 1:nrow(vuln_am_keys)) {
  species_range_raster <- rast(vuln_am_keys$file_nm[sp])

  # Temporary - exclude cells where probability < 0.05 and set remaining cells to 1 for presence
  species_range_raster[species_range_raster < 0.05] <- NA 
  species_range_raster[!is.na(species_range_raster)] <- 1

  # Get total species range area
  vuln_am_keys$total_species_area[sp] <- global(
    cellSize(species_range_raster, unit = "km") * !is.na(species_range_raster), "sum", na.rm = TRUE
    )$sum

  # Check species extent against farm extents to see if it's worth continuing
  pre_check <- lapply(fa_rasts, function(ras) {
      check1 <- !is.null(terra::intersect(ext(ras), ext(species_range_raster))) # do extents overlap?
      if (check1) {
        check2 <- crop(species_range_raster, ras) %>% # is any value != NA?
          minmax() %>% as.vector() %>% is.na() %>% any() %>% isFALSE()
      } else {check2 <- F}
      all(check1, check2) # if any are F ~ F
    }) %>% unlist()

  # For areas worth overlaying (passed pre-check)
  overlap_rasts_fa <- fa_rasts[pre_check]
  overlap_rasts_sp <- lapply(overlap_rasts_fa, function(ras) {
      sp_rast <- crop(species_range_raster, ras)
      sp_rast <- resample(sp_rast, ras, method = "bilinear")
      overlap_mask <- !is.na(sp_rast) & !is.na(ras)
      global(cellSize(overlap_mask, unit = "km") * overlap_mask, "sum", na.rm = TRUE)$sum
    })
  vuln_am_keys$farm_overlap_area[sp] <- overlap_rasts_sp %>% unlist() %>% sum()

  if (sp %in% as.integer(seq(0,nrow(vuln_am_keys), length.out = 21))) {
    print(paste0(sp, " of ", nrow(vuln_am_keys), " done, ",
                 round(100*sp/nrow(vuln_am_keys),0), "% finished at ", Sys.time()))
  }
}

vuln_am_keys <- vuln_am_keys %>% 
  mutate(overlap_percent = 100*farm_overlap_area/total_species_area)

qsave(vuln_am_keys, file.path(impacts_path, "vulnerability_areaoverlap_allspecies.qs"))

vuln_am_keys <- vuln_am_keys %>% 
  filter(overlap_percent > 0) %>% 
  arrange(-overlap_percent)

qsave(vuln_am_keys, file.path(impacts_path, "vulnerability_areaoverlap_nonzero.qs"))
```

In this very crude and liberal estimation, the maximum exposure of any species to farm nutrient inputs was `r round(max(vuln_am_keys$overlap_percent), 2)`%.

```{roverlap-plot}
xbreaks <- c(min(vuln_am_keys$overlap_percent/100), max(vuln_am_keys$overlap_percent/100)) %>% log()
xbreaks <- seq(xbreaks[1], xbreaks[2], length.out = 5)
xlabels <- exp(xbreaks)

scientific_10 = function(x) {
  ifelse(
    x==0, "0",
    parse(text = sub("e[+]?", " %*% 10^", scales::scientific_format()(x)))
  )
}

ggplot(vuln_am_keys, aes(x = log(overlap_percent/100), y = vuln, colour = taxon)) +
  geom_point() +
  theme_classic() +
  scale_x_continuous(breaks = xbreaks, labels = scientific_10(xlabels)) +
  labs(y = "Vulnerability", x = "% overlap with species range")
```