### 2025-06-18 08_relative_N_inputs.qmd

---
title: "08_relative_N_inputs"
format: html
editor: source
---

```{r setup}
library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
library(here)
library(terra)
library(sf)
library(qs)
library(ggplot2)
library(tidyterra)
library(cowplot)
library(rnaturalearth)

source("00_dirs.R")

overwrite <- F
```

# Base rasters for resolution

Create a blank raster with enough resolution to get farms in there (100m)

```{r base-raster}
base_rast <- rast(res = 0.1/111.32, extent = ext(-180, 180, -90, 90), crs = "EPSG:4326")

inset_boxes <- list(
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 72),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -44, -39.75)
)

inset_rasters <- list(
  CAN1 = crop(base_rast, ext(inset_boxes[["CAN1"]])),
  CAN2 = crop(base_rast, ext(inset_boxes[["CAN2"]])),
  EUR  = crop(base_rast, ext(inset_boxes[["EUR"]])),
  CHI  = crop(base_rast, ext(inset_boxes[["CHI"]])),
  AUS  = crop(base_rast, ext(inset_boxes[["AUS"]]))
)
```

# Raw background nitrogen

Data was downloaded from https://doi.org/10.5063/F1610XPS. Resolution is ~934 x 934 m in the Mollweide projection.

```{r raw-back-N}
back_N_rast <- file.path(input_Ndata_path, "nutrient_pollution_2013_raw.tif") %>% 
    rast()
back_N_rast
```

When extracting background N data, it was found that a lot of cells had no data. To try and fix this somewhat, I'm going to reproject the background N onto a slightly coarser grid.

```{r small-res-change}
back_N_rast <- back_N_rast %>% 
  project(rast(res = c(1000, 1000), extent = ext(back_N_rast), crs = crs(back_N_rast)),
          method  = "average")
back_N_rast
```

# Load modelled N data

```{r farm locations}
farm_locs <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread()
farm_sfs <- st_sf(farm_id = farm_locs$farm_id, geometry = farm_locs$geometry) %>% 
  st_transform(crs = crs(back_N_rast))
farm_rast <- farm_sfs %>% 
  rasterize(back_N_rast, field = "farm_id")
farm_rast
```

# Combine with background N

```{r}
# 2. Extract 3x3 neighborhood data for each farm point
# Get cell numbers for each farm point
farm_cells <- cellFromXY(back_N_rast, st_coordinates(farm_sfs))

# Function to get 5x5 neighborhood cell numbers
get_neighborhood_cells <- function(raster, center_cells) {
  # Get raster dimensions
  nrows <- nrow(raster)
  ncols <- ncol(raster)

  # Convert cell numbers to row/col indices
  rows <- rowFromCell(raster, center_cells)
  cols <- colFromCell(raster, center_cells)

  # Create all 25 neighbor offsets (including center) for 5x5 grid
  offsets <- expand.grid(row_offset = -2:2, col_offset = -2:2)

  # Initialize list to store results
  all_neighbors <- vector("list", length(center_cells))

  for(i in seq_along(center_cells)) {
    # Calculate neighbor row/col positions
    neighbor_rows <- rows[i] + offsets$row_offset
    neighbor_cols <- cols[i] + offsets$col_offset

    # Check bounds and filter valid cells
    valid_cells <- neighbor_rows >= 1 & neighbor_rows <= nrows & 
                   neighbor_cols >= 1 & neighbor_cols <= ncols

    # Convert back to cell numbers
    neighbor_cells <- cellFromRowCol(raster, 
                                   neighbor_rows[valid_cells], 
                                   neighbor_cols[valid_cells])

    # Create position labels for 5x5 grid
    row_names <- c("T2", "T1", "Crow", "B1", "B2")
    col_names <- c("L2", "L1", "Ccol", "R1", "R2")

    position_grid <- outer(row_names, col_names, paste, sep = "_")
    position_labels <- as.vector(position_grid)
    valid_positions <- position_labels[valid_cells]

    all_neighbors[[i]] <- data.frame(
      farm_id = farm_sfs$farm_id[i],
      cell_id = neighbor_cells,  # for overlap detection
      position = valid_positions,
      stringsAsFactors = FALSE
    )
  }
  return(do.call(rbind, all_neighbors))
}

# Get all neighborhood cells
neighborhood_data <- get_neighborhood_cells(back_N_rast, farm_cells) %>% 
  mutate(
    category = case_when(position == "Crow_Ccol" ~ "farm_direct",
                         position %in% c("T1_L1", "T1_Ccol", "T1_R1", "B1_L1", "B1_Ccol", "B1_R1", "Crow_L1", "Crow_R1") ~ "adjacent_1",
                         T ~ "adjacent_2"),
    category = factor(category, levels = c("farm_direct", "adjacent_1", "adjacent_2")),
    position = factor(position, levels = c(
      "T2_L2",   "T2_L1",   "T2_Ccol",   "T2_R1",   "T2_R2", 
      "T1_L2",   "T1_L1",   "T1_Ccol",   "T1_R1",   "T1_R2", 
      "Crow_L2", "Crow_L1", "Crow_Ccol", "Crow_R1", "Crow_R2", 
      "B1_L2",   "B1_L1",   "B1_Ccol",   "B1_R1",   "B1_R2", 
      "B2_L2",   "B2_L1",   "B2_Ccol",   "B2_R1",   "B2_R2"
    )))

# Extract nutrient values and combine with cell_ids
nutrient_values <- terra::extract(back_N_rast$nutrient_pollution_2013_raw, neighborhood_data$cell_id)
neighborhood_data <- cbind(neighborhood_data, nutrient_values) %>% 
  rename(np = nutrient_pollution_2013_raw)

# Fill gaps from original data with 8 neighbours
neighborhood_data_gaps <- neighborhood_data %>% filter(is.na(np))
for (i in 1:nrow(neighborhood_data_gaps)) {
  adj_cells <- adjacent(back_N_rast$nutrient_pollution_2013_raw, cells = neighborhood_data_gaps$cell_id[i], directions = 8)
  neighborhood_data_gaps$np[i] <- back_N_rast$nutrient_pollution_2013_raw[adj_cells[1,]] %>% 
    pull(nutrient_pollution_2013_raw) %>% 
    mean(na.rm = T)
}
neighborhood_data <- neighborhood_data %>% 
  filter(!is.na(np)) %>% 
  rbind(neighborhood_data_gaps)
neighborhood_data %>% filter(is.na(np)) %>% nrow() # 3221 still blank
neighborhood_data %>% filter(is.na(np) & position == "Crow_Ccol") %>% nrow() # 87 of those are direct farm cells

# Fill gaps from original data with 16 neighbours
neighborhood_data_gaps <- neighborhood_data %>% filter(is.na(np))
for (i in 1:nrow(neighborhood_data_gaps)) {
  adj_cells <- adjacent(back_N_rast$nutrient_pollution_2013_raw, cells = neighborhood_data_gaps$cell_id[i], directions = 16)
  neighborhood_data_gaps$np[i] <- back_N_rast$nutrient_pollution_2013_raw[adj_cells[1,]] %>% 
    pull(nutrient_pollution_2013_raw) %>% 
    mean(na.rm = T)
}
neighborhood_data <- neighborhood_data %>% 
  filter(!is.na(np)) %>% 
  rbind(neighborhood_data_gaps)
neighborhood_data %>% filter(is.na(np)) %>% nrow() # 4489 still blank
neighborhood_data %>% filter(is.na(np) & position == "Crow_Ccol") %>% nrow() # 149 of those are direct farm cells
```

```{r summary}
# Summary statistics
summary_stats <- neighborhood_data %>%
  group_by(farm_id, category) %>%
  reframe(
    mean = mean(np, na.rm = T),
    min = min(np, na.rm = T),
    max = max(np, na.rm = T),
    cells = n()
  )

# Maybe use adjacent farms to fill gaps?

ggplot(neighborhood_data, aes(x = category, y = log(np), fill = position)) +
  geom_boxplot()
```

# Plot mean nutrient pollution

```{r}
farm_N <- summary_stats %>% 
  filter(category == "farm_direct")

farm_N_sfs <- st_sf(farm_N = farm_N$mean, geometry = farm_locs$geometry) %>% 
  st_transform(crs = crs(back_N_rast))

plot(farm_N_sfs)
```