### 2025-07-08 08_relative_N_inputs.qmd
All changes

---
title: "08_relative_N_inputs"
author: Tormey Reimer
date: today
editor: source
---

# Introduction

The purpose of this markdown is to:

1. Import and clean up background nitrogen data
2. Overlay background nitrogen and Atlantic salmon farm locations
3. Compare annual inputs from the different sources

```{r setup}
#| code-summary: R setup

library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
library(here)
library(terra)
library(sf)
library(qs)
library(ggplot2)
library(tidyterra)
library(cowplot)
library(rnaturalearth)
library(purrr)

source("00_dirs.R")
feed_pal <- c("past" = "#E41A1C", "reference" = "#377EB8", "future" = "#4DAF4A")

prettyplot <- function() {
  theme_classic() +
    theme(legend.position = "none",
          text = element_text(family = "serif", size = 12, colour = "black"),
          axis.title.x = element_text(vjust = 0.5),
          axis.title.y = element_text(hjust = 0.5))
}
```

# Base rasters for resolution

```{r base-raster}
#| code-summary: Create some blank rasters with parameters to be used going forward

inset_boxes <- list(
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 72),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -44, -39.75)
)

base_wgs84 <- rast(res = c(0.008333333, 0.008333333), extent = ext(c(-180, 180, -90, 90)), crs = "EPSG:4326")
base_molle <- rast(res = c(1000, 1000), extent = ext(c(-18040096, 18040096, -9020048, 9020048)), crs = "ESRI:54009")

wgs84_insets <- inset_boxes %>% 
  lapply(function(bx) {crop(base_wgs84, ext(bx))})

molle_insets <- wgs84_insets %>% 
  lapply(function(rs) {project(rs, crs(base_molle))}) %>% 
  lapply(function(rs) {ext(rs)}) %>% 
  lapply(function(ex) {crop(base_molle, ex)})
```

# Raw background nitrogen

Data for nutrient runoff from agriculture fertilisers and pesticides was downloaded from @halpern_recent_2019-1 [dataset here: @halpern_recent_2019]. The data is in annual tonnes N.

```{r raw-back-N-halpern}
#| code-summary: Import data from @halpern_recent_2019 and crop/reproject for the 5 insets being used
#| eval: false

halpern_N_rast <- file.path(input_Ndata_path, "halpern_2019", "nutrient_pollution_2013_raw.tif") %>%
  rast()

halpern_insets <- molle_insets %>% 
  lapply(function(ins) {project(halpern_N_rast, ins, method = "cubic")})
```

Data for inputs of human sewage in coastal ecosystems was sourced from @tuholske_global_2021 [dataset here: @tuholske_mapping_2021]. The data is in annual g N.

```{r raw-back-N-tuholske}
#| code-summary: Import data from @tuholske_global_2021 and crop/reproject for the 5 insets being used
#| eval: false

tuholske_N_rast <- file.path(input_Ndata_path, "tuholske_2021", "global_effluent_2015_tot_N.tif") %>%
  rast() 
tuholske_N_rast <- tuholske_N_rast * 10^-6 # Change from g to t

# Change to the Molleweide projection with insets
tuholske_insets <- molle_insets %>% 
  lapply(function(ins) {project(tuholske_N_rast, ins, method = "cubic")})

# The halpern data has NA values for land but the tuholske data does not, so mask
tuholske_insets <- map2(tuholske_insets, halpern_insets, function(rast, mask) {
  mask(rast, mask)
})
```

```{r save-back-N}
#| code-summary: Combine and save background N data
#| eval: false

background_N_insets <- map2(
  halpern_insets, 
  tuholske_insets, 
  function(halpern, tuholske) {
    halpern + tuholske
  })

map2(names(halpern_insets), background_N_insets, function(nm, rast) {
  writeRaster(rast, 
              file.path(input_Ndata_path, str_c("background_N_inset_", nm, ".tif")), 
              overwrite = T)
})
```

The code above was only run once, background N rasters can now be loaded.

```{r load-background-data}
#| code-summary: Load previously saved inset rasters

background_N_insets <- lapply(names(inset_boxes), function(nm) {
  file.path(input_Ndata_path, str_c("background_N_inset_", nm, ".tif")) %>% rast()
})
```

# Combine modelled N data with background N

```{r farm locations}
#| code-summary: Bring in modelled salmon inputs data

farm_locs <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread()
farm_sfs <- st_sf(farm_id = farm_locs$farm_id, geometry = farm_locs$geometry) %>% 
  st_transform(crs = crs(base_molle))
farm_id_insets <- background_N_insets %>% 
  lapply(function(ins) {rasterize(farm_sfs, ins, field = "farm_id")})
```

```{r matching cells}
#| code-summary: Get background nitrogen data for farm cells

farm_cells <- map(background_N_insets, function(back_N_rast) {
  # Get cell numbers for each farm point
  farm_cells <- cellFromXY(back_N_rast, st_coordinates(farm_sfs))
  farm_cells <- farm_locs %>% 
    mutate(cell = farm_cells) %>% 
    filter(!is.na(cell))

  # Extract nutrient values and combine with cell_ids
  farm_cells$back_N <- terra::extract(back_N_rast$nutrient_pollution_2013_raw,
                                      farm_cells$cell)$nutrient_pollution_2013_raw

  # Fill gaps from original data with 8 neighbours
  farm_cells_gaps <- farm_cells %>% filter(is.na(back_N))
  for (i in 1:nrow(farm_cells_gaps)) {
    adj_cells <- adjacent(back_N_rast$nutrient_pollution_2013_raw, 
                          cells = farm_cells_gaps$cell[i], directions = 8)
    adj_cells_N <- back_N_rast$nutrient_pollution_2013_raw[adj_cells[1,]] %>% 
      pull(nutrient_pollution_2013_raw) %>% 
      mean(na.rm = T)
    farm_cells_gaps$back_N[i] <- adj_cells_N
  }
  farm_cells <- farm_cells %>% 
    filter(!is.na(back_N)) %>% 
    rbind(farm_cells_gaps)

  # Fill gaps from original data with 16 neighbours
  farm_cells_gaps <- farm_cells %>% filter(is.na(back_N))
  if (nrow(farm_cells_gaps) != 0) {
      for (i in 1:nrow(farm_cells_gaps)) {
      adj_cells <- adjacent(back_N_rast$nutrient_pollution_2013_raw, 
                            cells = farm_cells_gaps$cell[i], directions = 16)
      adj_cells_N <- back_N_rast$nutrient_pollution_2013_raw[adj_cells[1,]] %>% 
        pull(nutrient_pollution_2013_raw) %>% 
        mean(na.rm = T)
      farm_cells_gaps$back_N[i] <- adj_cells_N
    }
    farm_cells <- farm_cells %>% 
      filter(!is.na(back_N)) %>% 
      rbind(farm_cells_gaps)
  }

  default_val <- exp(median(log(farm_cells$back_N), na.rm = T))
  farm_cells$back_N[is.na(farm_cells$back_N)] <- default_val
  farm_cells$back_N[farm_cells$back_N < 0] <- default_val
  farm_cells
}) %>% 
  bind_rows()
```

```{r combine with N input from farms}

N_input <- output_cohorts_data_path %>% list.files(full.names = T) %>% 
  str_subset("allfarms") %>% 
  str_subset("P_excr") %>% 
  qs::qread() %>% 
  mutate(mean = mean * 6.25) %>% 
  group_by(farm_ID, feed) %>% 
  reframe(mean = sum(mean) * 10^-6) %>% 
  pivot_wider(names_from = feed, values_from = mean)

N_comp <- merge(farm_cells, N_input, by.y = "farm_ID", by.x = "farm_id", all = T) %>% 
  select(-cell) %>% 
  mutate(reference = reference/back_N,
         past = past/back_N,
         future = future/back_N) %>% 
  pivot_longer(names_to = "feed", names_transform = list(feed = as.factor),
               values_to = "ratio", cols = c("reference", "past", "future")) %>% 
  mutate(feed = factor(feed, levels = c("reference", "past", "future")))
```

# Analyse

```{r summary-plots}

N_comp %>%
  group_by(country, farm_id, feed) %>% 
  reframe(mean = mean(ratio, na.rm = T)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(past = (past-reference)/reference,
         future = (future-reference)/reference) %>% 
  select(-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "mean_ratio", cols = c(past, future)) %>% 
  ggplot(aes(x = country, y = mean_ratio, fill = feed)) +
  geom_boxplot() +
  prettyplot() +
  scale_y_continuous(breaks = seq(0,1,0.2), labels = seq(0,100,20), limits = c(0,0.6)) +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Mean change in N input ratio", x = "Country")

N_comp %>% 
  filter(log(ratio) < 20) %>%
  ggplot(aes(x = country, y = log(ratio), fill = feed)) +
  geom_boxplot() +
  prettyplot() +
  # scale_y_continuous(breaks = seq(0,1,0.2), labels = seq(0,100,20), limits = c(0,0.6)) +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Mean change in N input ratio", x = "Country")
```