### 2025-07-10 06_run_atlantic_salmon_model.qmd

---
title: "Individual Mariculture Modeling Analysis"
format: html
editor: source
---

# Introduction

This document runs individual mariculture modeling analyses for Atlantic salmon farms. It processes farm location data, species parameters, and conducts sensitivity analyses to understand the impact of various parameters on fish and farm growth measures.

```{r setup}
#| warning: false
#| message: false

library(arrow)
library(sf)
library(dplyr)
library(tidyr)
library(terra)
library(magrittr)
library(furrr)
library(future)
library(tictoc)
library(ggplot2)
library(fs)
library(conflicted)
library(stringr)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)
```

```{r filenames}
# Set up parallel processing
# plan(multisession, workers = parallelly::availableCores()-1)

# Filenames
species_params_excel <- c(file = file.path(input_species_param_path, "Species.xlsx"), sheet = "Atlantic salmon")
pop_params_excel <- c(file = file.path(input_species_param_path, "Population.xlsx"))
farm_locations_parquet <- file.path(input_farm_coords_path, "farm_coords.parquet")
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
farm_geometry_file <- file.path(output_farm_data_path, "farm_geometry.qs")
farm_ts_data_file <- file.path(output_farm_data_path, "farm_ts_data.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
farm_harvest_file <- file.path(output_farm_data_path, "farm_harvest_size.qs")
```

# Data for targets pipelines

Much of the actual analysis is run through targets pipelines. Therefore, we need to make sure that the files going into those pipelines are correct and up to date.

Load and process farm coordinate data with appropriate timing parameters for Northern and Southern hemisphere farms.

```{r farm-coordinates}
times_N <- c("t_start" = 121, "t_end" = 121+547, "dt" = 1)
times_S <- c("t_start" = 274, "t_end" = 274+547, "dt" = 1)

farm_coords <- farm_locations_parquet %>% 
  read_parquet() %>% 
  mutate(t_start = case_when(lat > 0 ~ times_N['t_start'], TRUE ~ times_S['t_start']), 
          t_end = case_when(lat > 0 ~ times_N['t_end'], TRUE ~ times_S['t_end']),
          t_start = unname(t_start),
          t_end = unname(t_end))

qsave(farm_coords, farm_coords_file)

# Also save geometry for later
file.path(input_farm_coords_path, "atlantic_salmon_locations_w_temps.qs") %>% 
  qread() %>% 
  dplyr::filter(day == "day_1") %>% 
  dplyr::select(farm_id, geometry, country) %>% 
  qsave(farm_geometry_file)
```

## Farm Time Series Data

Process Sea Surface Temperature (SST) data for each farm location.

```{r farm-ts-data}
farms_to_omit <- qread(sprintf(file.path(input_farm_coords_path, "%s_farms_to_omit.qs"), this_species))
farm_SST_data <- read_parquet(file.path(input_farm_sst_path, "farm_SST_extracted.parquet"))
farm_IDs <- farm_SST_data %>%
  filter(!farm_id %in% farms_to_omit) %>%
  distinct(farm_id) %>%
  pull(farm_id)

farm_ts_data <- farm_SST_data %>%
  rename(farm_ID = farm_id) %>% 
  select(c(farm_ID, day, temp_c)) %>%
  mutate(day = str_split_i(day, "day_", 2) %>% as.integer())

qsave(farm_ts_data, farm_ts_data_file)
```

# Species and population parameters

Load species-specific parameters from Excel file.

```{r species-parameters}
species_params <- readxl::read_excel(
  path = species_params_excel["file"], 
  sheet = species_params_excel["sheet"]
)
vals <- species_params$Value
names(vals) <- species_params$Quantity
species_params <- vals[!is.na(vals)]
qsave(species_params, species_params_file)
```

Load population-specific parameters.

```{r population-parameters}
pop_params <- readxl::read_excel(path = pop_params_excel["file"])
vals <- pop_params$Value
names(vals) <- pop_params$Quantity
pop_params <- vals[!is.na(vals)]
qsave(pop_params, pop_params_file)
```

# Farm Harvest Size Calculations

Calculations of expected harvest sizes for each farm (using default feed and no Monte-Carlo variation) are done in the targets pipeline.

```{r farm-harvest-size}
Sys.setenv(TAR_PROJECT = "project_farmruns")

targets::tar_make(harvest_size)
targets::tar_load(harvest_size)
qsave(harvest_size, farm_harvest_file)

mean(harvest_size$weight)
min(harvest_size$weight)
max(harvest_size$weight)

ggplot(harvest_size, aes(x = weight)) +
  geom_histogram(fill = "salmon", alpha = 0.75, colour = "black")
```

# Sensitivity analysis

Run the sensitivity runs using the targets pipeline. 
Note that this can take a long time (hours), as the species parameters (19) are run for each farm (2721) and factor (3) and the population parameters (6) are run for a smaller number of farms (271) and factor with a simulated population of 500 fish.

```{r sensitivity-run}
Sys.setenv(TAR_PROJECT = "project_sensitivities")

targets::tar_validate()
targets::tar_visnetwork(targets_only = T)
targets::tar_outdated()
targets::tar_make(
  seconds_meta_append = 120
  )
```

## Process sensitivity results

Combine and visualise sensitivity analysis results.

```{r sensitivity-results}
targets::tar_load(sens_results_spec)
targets::tar_load(sens_results_pop)

sens_results <- rbind(
  sens_results_pop,
  sens_results_spec
)

sens_measures <- levels(sens_results$measure)
sens_results_files <- file.path(output_sens_data_path, paste0("sens_results_", sens_measures, ".qs"))
sens_results_figfiles <- file.path(output_sens_data_path, paste0("sens_plot_", sens_measures, ".qs"))
sens_results_figfiles2 <- file.path(output_sens_data_path, paste0("sens_plot_", sens_measures, ".png"))

for (sm in seq_along(sens_measures)) {
  sens_results %>% 
    filter(measure == sens_measures[sm]) %>% 
    qsave(sens_results_files[sm])

  p <- sens_results %>% 
    filter(measure == sens_measures[sm]) %>% 
    ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens)) +
    geom_col(fill = "salmon", alpha = 0.65, colour = "black") +
    geom_errorbar(width = 0.5) +
    coord_flip() +
    theme_classic()
  qsave(p, sens_results_figfiles[sm])
  ggsave(sens_results_figfiles2[sm])
}
```

# Run farms

```{r run-farms-pipeline}
Sys.setenv(TAR_PROJECT = "project_farmruns")

targets::tar_validate()
targets::tar_visnetwork(targets_only = T)
targets::tar_outdated()
targets::tar_make(seconds_meta_append = 90)
```

```{r farm-run-results}
targets::tar_load(cohort_results)
```