### 2025-07-23 manuscript.qmd
Docs fixup and render

---
title: "Local to global N input modelling"
# Author roles: "writing – original draft", "writing – review & editing", "formal analysis", "funding acquisition", "conceptualization", "data curation", "investigation", "methodology", "project administration", "resources", "software", "supervision", "validation", "visualization"

bibliography: ["refs.bib", "manual-refs.bib"]

format: 
  html:
    theme: simplex
    css: "extra.css"
    code-fold: true
    code-overflow: wrap
    toc: true
    toc-expand: true
    toc-location: left
    lang: en-GB
    grid:
      sidebar-width: 250px
      body-width: 1200px
      margin-width: 250px
      gutter-width: 2em

execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true
  freeze: auto

knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 8.5

editor: source
---

```{r setup}
#| include: false
#| eval: true
#| cache: false
#| message: false

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate", "worrms", "aquamapsdata") %>% sort()

here("00_model_functions.R") %>% source()

remove_unit("gN")
install_unit("gN", "6.25 g") # 1/6.25
```

```{r global-variables}
#| include: false

source(here("00_model_functions.R"))
source(here("00_dirs.R"))

# Filenames
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
sens_results_files <- file.path(output_sens_data_path) %>% list.files(full.names = T) %>% str_subset("results")

species_params <- qread(species_params_file)
pop_params <- qread(pop_params_file)

prettyplot <- function() {
  theme_classic() +
    theme(legend.position = "none",
          text = element_text(family = "serif", size = 12, colour = "black"),
          axis.title.x = element_text(vjust = 0.5),
          axis.title.y = element_text(hjust = 0.5))
}

feed_pal <- c("past" = "#E41A1C", "reference" = "#377EB8", "future" = "#4DAF4A")  
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", "P" = "darkred", 
  "carbohydrates" = "steelblue", "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", "L" = "darkgoldenrod1"
)
```

# Methods
### Feeding rate

To detail here:

* Functional response to temperature (species-specific)
* Overfeeding parameters and how they come into play
* Deminstration of how feeding response and actual food given vary with temperature?

```{r fig-functional-response-to-temperature}
#| eval: false
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

df <- data.frame(water_temp = seq(-5, 26, 0.25))
df$feeding <- sapply(FUN = feeding_rate, X = df$water_temp, species_params = species_params)
df <- df %>% 
mutate(ing_pot = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * feeding,
       food_prov = NA)
for (i in 1:nrow(df)) {
  df$food_prov[i] <- food_prov_rate(pop_params = pop_params, rel_feeding = df$feeding[i], ing_pot = df$ing_pot[i], 
                                    ing_pot_10 = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * 0.1)
}
df$ing_pot <- df$ing_pot/maxna(df$ing_pot)
df$food_prov <- df$food_prov/maxna(df$ing_pot)

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  prettyplot() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

## Farm temperature forcings

```{r farm info}
#| code-summary: Get general farm info

farms_to_omit <- file.path(input_farm_coords_path, "atlantic_salmon_farms_to_omit.qs") %>% qs::qread()
farm_ts_data <- file.path(output_farm_data_path, "farm_ts_data.qs") %>% qs::qread()
farm_coords <- file.path(output_farm_data_path, "farm_coords.qs") %>% qs::qread() %>% 
    mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N") %>% as.factor())

farms_geometry <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread() %>% 
  rename(farm_ID = farm_id)

sorted_countries <- farm_coords %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  group_by(country) %>% 
  reframe(av_lat = mean(lat)) %>% 
  arrange(-av_lat) %>% 
  pull(country) %>% 
  unique()

farms_geometry <- farms_geometry %>% 
  mutate(country = factor(country, levels = sorted_countries))
```

```{r farm temperature data}
#| code-summary: Get the farm temperature forcings

farm_temp_means <- farm_ts_data %>% 
  group_by(farm_ID) %>% 
  reframe(mean_temp = mean(temp_c)) %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  dplyr::filter(!farm_ID %in% farms_to_omit)

overall_mean <- mean(farm_temp_means$mean_temp)
overall_sd <- sd(farm_temp_means$mean_temp)

rm(farm_ts_data)
```

Mean daily temperatures at the farms globally ranged from  from `r round(min(farm_temp_means$mean_temp), 1)`$^{\circ}$C to `r round(max(farm_temp_means$mean_temp), 1)`$^{\circ}$C, with an overall mean of `r round(overall_mean, 1)`$^{\circ}$C ($\pm$ `r round(overall_sd, 1)`$^{\circ}$C SD).
Farms with a mean temperature of $\leq$6$^\circ$C were excluded (`r as.integer(length(farms_to_omit))` farms). The total number of farms included in the analysis was therefore `r nrow(farm_temp_means)`.

```{r fig-farm-mean-temps}
#| fig-cap: |
#|   Potential options for showing data spatially. Top left: Mean temperature by latitude across countries. Top right: Frequency distribution of mean temperatures by hemisphere. Bottom: global distribution of mean temperatures.
#| fig-height: 9
#| fig-width: 9.5

farm_temp_robinson <- st_sf(farm_temp_means) %>% 
  st_transform(crs = "+proj=robin")

p_bigmap_robinson +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25, barheight = 1.5
  ))

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  prettyplot() +
  theme(legend.position = "top", legend.title = element_blank())

p3 <- farm_temp_means %>% 
  mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N")) %>% 
  ggplot(aes(x = mean_temp, fill = hemisphere)) +
  geom_histogram(position = "identity", colour = "black", alpha = 0.5, binwidth = 0.5) +
  labs(x = expression("Mean temperature ("*degree*"C)"), y = "Frequency") +
  prettyplot() +
  theme(legend.position = "top", legend.title = element_blank())

(p2 + p3) / p_big_map + plot_layout(heights = c(1, 1.5))
```

```{r patchwork-map-mercator}
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

p_bigmap_temperature <- p_bigmap_mercator +
  geom_sf(data = farm_temp_means, aes(color = mean_temp, geometry = geometry), size = 2) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

p_insets <- get_insets_mercator(p_bigmap_temperature)

grid_1 <- cowplot::plot_grid(
  p_insets[["CAN1"]], p_insets[["CAN2"]], 
  ncol = 2, rel_widths =  c(1, 1.35)
  )
grid_2 <- cowplot::plot_grid(p_Eur,  p_Aus,  nrow = 2, rel_heights = c(1, 1))
grid_3 <- cowplot::plot_grid(grid_2, p_Chi,  ncol = 2, rel_widths =  c(1, 1))

cowplot::plot_grid(
  grid_1, grid_3,
  nrow = 2,
  rel_heights = c(1, 1.5)        # relative row heights
)
```

## Modelled experimental feeds {#sec-feeds}

```{r feeds}
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>%
  qs::qread()

feed_summary <- map_dfr(names(feed_params), function(feed) {
  df <- feed_params[[feed]]
  map_dfr(c("Proteins", "Lipids", "Carbohydrates"), function(macro) {
    contrib <- df[[macro]]$proportion * df[[macro]]$macro
    total <- sum(contrib)
    digest <- sum(df[[macro]]$digest * contrib) / total
        tibble(
      feed = feed,
      macro = macro,
      total = total,
      digest = digest
    )
  })
}) %>%
  mutate(macro = factor(macro, levels = c("Proteins", "Lipids", "Carbohydrates"), labels = c("P", "L", "C")),
         feed = factor(feed, levels = c("reference", "past", "future")),
         total_r = total * 100 %>% round(1),
         digest_r = digest * 100 %>% round(1))
```

The reference feed was 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "reference", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "reference", macro == "C")$total, 1)`% carbohydrate, the past feed was 
`r round(100*filter(feed_summary, feed == "past", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "past", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "past", macro == "C")$total, 1)`% carbohydrate, and the future feed was 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "future", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "future", macro == "C")$total, 1)`% carbohydrate.

Ingredients are tracked seperately through the model with their own digestibility coefficients. 
Overall, protein digestibility was 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$digest, 1)`%,
`r round(100*filter(feed_summary, feed == "past", macro == "P")$digest, 1)`%, and 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$digest, 1)`% and carbohydrate digestibility was 
`r round(100*filter(feed_summary, feed == "reference", macro == "C")$digest, 1)`%,
`r round(100*filter(feed_summary, feed == "past", macro == "C")$digest, 1)`%, and 
`r round(100*filter(feed_summary, feed == "future", macro == "C")$digest, 1)`%
within the reference, past, and future feeds respectively. 
The digestibility of lipids was assumed to remain constant across all ingredients at `r round(100*filter(feed_summary, feed == "past", macro == "L")$digest, 1)`%.

```{r fig-feed-summary-ideas}
#| fig-height: 7.5
#| fig-cap: | 
#|   Ideas for how to present the differences between feeds. Top plot shows the protein (red), lipids (yellow) and carbohydrate (blue) composition, with the digestible portion shown darker and the undigestible portion shown lighter. Bottom plot shows the proportion of each feed composed of each types of ingredient.

p1 <- feed_summary %>%
  mutate(digest_total = digest * total) %>% 
  ggplot(aes(x = feed, y = total, fill = macro)) +
  geom_col(position = "dodge", alpha = 0.55, colour = "black") +
  geom_col(mapping = aes(y = digest_total), position = "dodge", alpha = 1, colour = "black") +
  scale_fill_manual(values = macro_pal) +
  scale_y_continuous(breaks = seq(0,0.5,0.1), labels = seq(0,50,10)) +
  scale_x_discrete(labels = c("reference" = "Reference", "past" = "Past", "future" = "Future")) +
  labs(x = "Feed", y = "Proportion (%)") +
  prettyplot()

feed_ings <- rbind(
  feed_params[[1]][[1]] %>% mutate(feed = as.factor("reference")), 
  feed_params[[2]][[1]] %>% mutate(feed = as.factor("past")), 
  feed_params[[3]][[1]] %>% mutate(feed = as.factor("future"))
) %>%
  rename(protein = macro) %>%
  mutate(
    category = case_when(
      str_detect(ingredient, "fish-oil") ~ "fish",
      str_detect(ingredient, "fishmeal") ~ "fish",
      str_detect(ingredient, "krill") ~ "other-marine",
      str_detect(ingredient, "soy") ~ "legume",
      str_detect(ingredient, "pea") ~ "legume",
      str_detect(ingredient, "guar") ~ "legume",
      str_detect(ingredient, "faba") ~ "legume",
      str_detect(ingredient, "corn") ~ "wheat-corn",
      str_detect(ingredient, "wheat") ~ "wheat-corn",
      str_detect(ingredient, "chicken") ~ "chicken",
      str_detect(ingredient, "bacteri") ~ "bact-fungi-micro",
      str_detect(ingredient, "yeast") ~ "bact-fungi-micro",
      str_detect(ingredient, "spirulina") ~ "bact-fungi-micro",
      str_detect(ingredient, "bsf") ~ "insect",
      T ~ "other-plant"
    ) %>% 
      factor(levels = c("fish", "other-marine", "chicken", "wheat-corn", "legume", "other-plant", "insect", "bact-fungi-micro"))
  )

p_ings <- feed_ings %>% 
  mutate(feed = factor(feed, levels = c("reference", "past", "future"), labels = c("Reference", "Past", "Future"))) %>% 
  group_by(feed, category) %>%
  reframe(proportion = sum(proportion)) %>% 
  ggplot(aes(x = "", y = proportion, fill = category)) +
  geom_col(width = 1, colour = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(
    name = "Ingredient category",
    values = c("fish" = "royalblue3", "other-marine" = "turquoise2", "chicken" = "hotpink1", "wheat-corn" = "moccasin", "legume" = "forestgreen", "other-plant" = "darkolivegreen1", "insect" = "firebrick3", "bact-fungi-micro" = "darkorange"), 
    labels = c("Fish meal or oil", "Other marine ingredients", "Chicken", "Wheat and corn", "Legumes", "Other plant ingredients", "Insects", "Bacteria, fungi or microalgae")) +
  facet_grid(cols = vars(feed)) +
  theme_void() +
  theme(
    text = element_text(family = "serif", size = 12, colour = "black"),
    legend.position = "bottom",
    legend.title = element_blank(), 
    strip.text = element_blank(),
    strip.text.x = element_text(size = 14)
  )

plot_grid(p1, p_ings, nrow = 2)
```

## Impacts of nutrient input

General thoughts is that a risk-assessment approach is most compelling:

* Hazard: Nutrient input. Present as a change between feeds (%?), relative to background levels (ratio?)
  * Possibly ask Camille how the quotas for Tasmanian salmon are calculated
  * Background levels are from [oceanhealthindex.org](https://knb.ecoinformatics.org/view/resource_map_doi:10.5063/F1610XPS) - double-check that it includes wastewater input AND agriculture - should include both - if they're not combined ask Ben because he has them combined somewhere
  * If this is too uncertain perhaps combine into categorical (oligotrophic/eutrophic) classification?
* Exposure: Species presence. Use IUCN/Aquamap data to estimate the % of each species range is occupied by farms
  * give each farm a 1km radius, check for overlaps
  * Gage already has the data to process the Aquamap data
* Vulnerability: species vulnerability to eutrophication + nutrient input [@butt_trait-based_2022; @ohara_code_2021]. Possibly also use IUCN threat levels for prioritisation?

# Results

* Parameter sensitivity results are in supplementary @sec-parameter-sensitivity

In general, data outputs comes in the following dimensions:

* Location -> temperature -> farm_ID
* Uneaten feed + excreted faeces = total inputs (g)
* Day of production (impacted by location)
* Main point is the difference between feeds (kg/t salmon).
* Also interesting to see if difference between feeds varies geographically (correlates with latitude, or maybe mean/median/max temperature?)

## Uneaten feed and faeces

```{r inputs data}
#| code-summary: Load and prep uneaten feed and faeces data (keep in g faeces/feed)

mass_inputs_ts <- find_read(data_analysis_path, "all_waste_inputs")
mass_inputs_end <- find_read(data_analysis_path, "sumtotal_waste_inputs.qs")
N_inputs_ts <- find_read(data_analysis_path, "all_N_inputs")
N_inputs_end <- find_read(data_analysis_path, "sumtotal_N_inputs")
```

```{r inputs stats through time}
#| eval: false
#| code-summary: Show how inputs (mass and N) change through time (across all farms)

mass_inputs_ts
N_inputs_ts

# Through time
inputs_stats_ts <- inputs_df %>%
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

inputs_stats_rel_ts <- inputs_rel_df %>% 
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

# Total (at production end)
inputs_stats_end <- inputs_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

inputs_stats_rel_end <- inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

inputs_stats_rel_end_diff <- inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(future = (future-reference)/reference,
         past = (past-reference)/reference) %>% 
  select(-reference) %>% 
  pivot_longer(cols = c(past, future), names_to = "feed", values_to = "diff") %>% 
  mutate(feed = factor(feed, levels = c("past", "future"))) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(diff),
          max = maxna(diff),
          sd = sdna(diff),
          mean = meanna(diff))

# Pull key values (uneaten feed)
u_stmin1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(min)
u_stmax1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(max)
u_stmean1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "uneat") %>% pull(mean)
u_stmin2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(min)
u_stmax2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(max)
u_stmean2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "uneat") %>% pull(mean)
u_stmin3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(min)
u_stmax3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(max)
u_stmean3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "uneat") %>% pull(mean)

# Convert past and future values to % change values
u_stmin2 <- (100*u_stmin2) %>% round(2)
u_stmax2 <- (100*u_stmax2) %>% round(2)
u_stmean2 <- (100*u_stmean2) %>% round(2)
u_stmin3 <- (100*u_stmin3) %>% round(2)
u_stmax3 <- (100*u_stmax3) %>% round(2)
u_stmean3 <- (100*u_stmean3) %>% round(2)

u_stmin1 <- u_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmax1 <- u_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmean1 <- u_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)

# Pull key values (excreted faeces)
e_stmin1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(min)
e_stmax1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(max)
e_stmean1 <- filter(inputs_stats_rel_end, feed == "reference", measure == "excr") %>% pull(mean)
e_stmin2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(min)
e_stmax2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(max)
e_stmean2 <- filter(inputs_stats_rel_end_diff, feed == "past", measure == "excr") %>% pull(mean)
e_stmin3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(min)
e_stmax3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(max)
e_stmean3 <- filter(inputs_stats_rel_end_diff, feed == "future", measure == "excr") %>% pull(mean)

# Convert past and future values to % change values
e_stmin2 <- (100*e_stmin2) %>% round(2)
e_stmax2 <- (100*e_stmax2) %>% round(2)
e_stmean2 <- (100*e_stmean2) %>% round(2)
e_stmin3 <- (100*e_stmin3) %>% round(2)
e_stmax3 <- (100*e_stmax3) %>% round(2)
e_stmean3 <- (100*e_stmean3) %>% round(2)

e_stmin1 <- e_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
e_stmax1 <- e_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
e_stmean1 <- e_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
```

```{r waste inputs stats total}
#| code-summary: Show how waste inputs change in total (at the end of a production period) across all farms

mass_inputs_sep <- mass_inputs_end %>% 
  mutate(waste_per_biomass = set_units(total/biomass_total, "g kg_fish-1")) %>% 
  dplyr::select(-c(total, biomass_total)) %>% 
  pivot_wider(names_from = measure, values_from = waste_per_biomass) %>% 
  mutate(total = excr+uneat) %>% 
  pivot_longer(names_to = "measure", values_to = "waste_per_biomass", cols = c(total, excr, uneat), names_transform = list(measure = as.factor)) %>% 
  group_by(feed, measure) %>% 
  reframe(mean_wpb = meanna(waste_per_biomass),
          mean_wpb = sdna(waste_per_biomass) %>% set_units("g kg_fish-1"),
          min_wpb = minna(waste_per_biomass),
          max_wpb = maxna(waste_per_biomass))

twi_min <- mass_inputs_sep$min_wpb[mass_inputs_sep$measure == "total"] %>% round(2)
twi_max <- mass_inputs_sep$max_wpb[mass_inputs_sep$measure == "total"] %>% round(2)
twi_mean <- mass_inputs_sep$mean_wpb[mass_inputs_sep$measure == "total"] %>% round(2)

# There were only very small differences in total uneaten feed between the different feeds, with the past feed increasing uneaten feed by `r u_stmin2`-`r u_stmax2`%  (mean of `r u_stmean2`%) and the future feed increasing uneaten feed by `r u_stmin3`-`r u_stmax3`%  (mean of `r u_stmean3`%).

# * Total excreted faeces across a cohort fed the reference feed ranged from `r e_stmin1` to `r e_stmax1` g kg$^{-1}$ fish, with a mean of `r e_stmean1` g kg$^{-1}$ fish. 
# * Past feed decreased excreted faeces by `r -1*e_stmin2`-`r -1*e_stmax2`%  (mean of `r -1*e_stmean2`%) 
# * Future feed increased excreted faeces by `r e_stmin3`-`r e_stmax3`%  (mean of `r e_stmean3`%)
```

* Total waste inputs across a single production period of fish fed the reference feed ranged from `r twi_min[1]` to `r twi_max[1]` g kg$^{-1}$ fish, with a mean of `r twi_mean[1]` g kg$^{-1}$ fish.
* Total waste inputs across a single production period of fish fed the past feed ranged from `r twi_min[2]` to `r twi_max[2]` g kg$^{-1}$ fish, with a mean of `r twi_mean[2]` g kg$^{-1}$ fish.
* Total waste inputs across a single production period of fish fed the future feed ranged from `r twi_min[3]` to `r twi_max[3]` g kg$^{-1}$ fish, with a mean of `r twi_mean[3]` g kg$^{-1}$ fish.

```{r fig-waste-inputs-total}
#| fig-cap: | 
#|   Total waste intputs (uneaten feed + excreted faeces) across the whole production period at each farm location.
#| fig-width: 9.5

mass_inputs_sep_geometry <- mass_inputs_end %>% 
  mutate(waste_per_biomass = set_units(total/biomass_total, "g kg_fish-1")) %>% 
  dplyr::select(-c(total, biomass_total)) %>% 
  pivot_wider(names_from = measure, values_from = waste_per_biomass) %>% 
  mutate(total = excr+uneat) %>% 
  pivot_longer(names_to = "measure", values_to = "waste_per_biomass", cols = c(total, excr, uneat), names_transform = list(measure = as.factor)) %>% 
  group_by(farm_ID, feed, measure) %>% 
  reframe(mean_wpb = meanna(waste_per_biomass),
          sd_wpb = sdna(waste_per_biomass) %>% set_units("g kg_fish-1"),
          min_wpb = minna(waste_per_biomass),
          max_wpb = maxna(waste_per_biomass)) %>% 
  merge(farms_geometry, by = "farm_ID")

mass_inputs_robinson <- mass_inputs_sep_geometry %>% 
  filter(feed == "reference" & measure == "total") %>% 
  mutate(mean_wpb = mean_wpb %>% drop_units()) %>% 
  st_sf() %>% 
  st_transform(crs = "+proj=robin")

ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = mass_inputs_robinson, aes(color = mean_wpb), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = expression("Mean waste inputs (g g fish"^-1*")"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()
```

```{r patchwork-waste-map-mercator}
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

mass_inputs_patch <- mass_inputs_sep_geometry %>% 
  filter(feed == "reference" & measure == "total") %>% 
  mutate(mean_wpb = mean_wpb %>% drop_units())

p1 <- ggplot() +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = mass_inputs_patch, aes(color = mean_wpb, geometry = geometry), size = 2) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = expression("Mean waste inputs (g g fish"^-1*")"),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1.5,
    barwidth = 1.5, barheight = 20,
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

p_Can1 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN1"]][["xlims"]], 
           ylim = inset_boxes_sm[["CAN1"]][["ylims"]])  +
  draw_label("A", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN1"]][["labx"]], 
             y = inset_boxes_sm[["CAN1"]][["laby"]], 
             hjust = -1, vjust = 0.5)
p_Can2 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN2"]][["xlims"]], 
           ylim = inset_boxes_sm[["CAN2"]][["ylims"]])  +
  draw_label("B", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN2"]][["labx"]], 
             y = inset_boxes_sm[["CAN2"]][["laby"]], 
             hjust = 0.5, vjust = 0.5)
p_Eur <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["EUR"]][["xlims"]], 
           ylim = inset_boxes_sm[["EUR"]][["ylims"]])  +
  draw_label("C", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["EUR"]][["labx"]], 
             y = inset_boxes_sm[["EUR"]][["laby"]], 
             hjust = -1, vjust = 0.5)
p_Chi <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
           ylim = inset_boxes_sm[["CHI"]][["ylims"]])  +
  theme(legend.position = "right") +
  draw_label("D", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CHI"]][["labx"]], 
             y = inset_boxes_sm[["CHI"]][["laby"]], 
             hjust = 0, vjust = 0)
p_Aus <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["AUS"]][["xlims"]], 
           ylim = inset_boxes_sm[["AUS"]][["ylims"]])  +
  draw_label("E", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["AUS"]][["labx"]], 
             y = inset_boxes_sm[["AUS"]][["laby"]], 
             hjust = 0, vjust = 0.5)

grid_1 <- cowplot::plot_grid(p_Can1, p_Can2, ncol = 2, rel_widths =  c(1, 1.35))
grid_2 <- cowplot::plot_grid(p_Eur,  p_Aus,  nrow = 2, rel_heights = c(1, 1))
grid_3 <- cowplot::plot_grid(grid_2, p_Chi,  ncol = 2, rel_widths =  c(1, 1))

cowplot::plot_grid(
  grid_1, grid_3,
  nrow = 2,
  rel_heights = c(1, 1.5)        # relative row heights
)
```

```{r fig-total-excreted-perday}
#| eval: false
#| fig-cap: |
#|   Feed left uneaten when fish are fed the reference (blue), past (red) or future (green) feed. Countries are arranged by descending mean latitude.

excreted_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(excr_biom_mean = excr_biom_mean %>% set_units("g g_fish-1 d-1") %>% set_units("g kg_fish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = excr_biom_mean, fill = feed)) +
  geom_boxplot() +
  prettyplot() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Excreted faeces", x = "Country")
```

```{r fig-total-excr-diff_perc, class.chunk = "aside"}
#| eval: false
#| fig-cap: |
#|   Mean % difference in excreted faeces of the two experimental feeds (compared to the reference feed) with increasing overall excretion.

df <- excreted_total %>% 
  select(-contains(c("sd", "biom"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kg_fish-1 d-1"),
         diff_perc = drop_units(diff/reference)) %>% 
    merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t - t_start + 1)

summ <- df %>%
  group_by(hemisphere, feed) %>% 
  reframe(mean = 100*meanna(diff_perc) %>% round(3),
          min = 100*meanna(diff_perc) %>% round(3),
          max = 100*maxna(diff_perc) %>% round(3))

df %>% 
  ggplot(aes(x = reference, y = diff_perc, colour = hemisphere)) +
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(~feed) +
  prettyplot()

# The difference in total excretion for the past feed ranged from 
# `r summ$min[summ$feed == "past"]`-`r summ$max[summ$feed == "past"]`% across all farms globally, and for the future feed ranged from 
# `r summ$min[summ$feed == "future"]`-`r summ$max[summ$feed == "future"]`%. 
```

```{r fig-total-excr-diff_perc_2, class.chunk = "aside"}
#| eval: false
#| fig-cap: |
#|   Mean % difference in excreted faeces of the two experimental feeds (compared to the reference feed) over time in one production cycle.

df %>% 
  group_by(prod_day, feed, hemisphere) %>% 
  reframe(diff_perc_mean = meanna(diff_perc),
          diff_perc_sd = sdna(diff_perc)) %>% 
  ggplot(aes(x = prod_day, y = diff_perc_mean, ymin = diff_perc_mean-diff_perc_sd, ymax = diff_perc_mean+diff_perc_sd, 
             colour = as.factor(hemisphere), fill = as.factor(hemisphere))) +
  geom_line() + 
  geom_ribbon(alpha = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(~feed) +
  prettyplot()
```

```{r fig-total-excr-act_diff}
#| eval: false
#| fig-width: 6.25
#| fig-height: 5.25
#| fig-cap: Mean actual difference in excreted faeces of the two experimental feeds (compared to the reference feed) over time in one production cycle.

df <- excreted_total %>% 
  select(-contains(c("sd", "excr_mean", "biomass"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_biom_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kg_fish-1 d-1"),
         diff = set_units(diff, "kg t_fish-1 d-1")) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t-t_start+1)

summ <- df %>%
  group_by(hemisphere, feed) %>% 
  reframe(mean = meanna(diff) %>% round(3),
          min = meanna(diff) %>% round(3),
          max = maxna(diff) %>% round(3))

df %>% 
  group_by(hemisphere, feed, prod_day) %>% 
  reframe(mean_diff = meanna(diff),
          sd_diff = sdna(diff) %>% set_units("kg t_fish-1 d-1")) %>% 
  ggplot(aes(x = prod_day, y = mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff, colour = hemisphere, fill = hemisphere)) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(~feed) +
  prettyplot()
```

## Total nutrient inputs

```{r protein raw data}
#| code-summary: Load and prep macronutrients data (keep in g faeces/feed)

N_inputs_df <- rbind(
  farmrun_files %>% 
    purrr::map(function(f) {qs::qread(f)[["P_excr_stat"]]}) %>% 
    bind_rows() %>% 
    mutate(measure = "excr"),
  farmrun_files %>% 
    purrr::map(function(f) {qs::qread(f)[["P_uneat_stat"]]}) %>% 
    bind_rows() %>% 
    mutate(measure = "uneat")
) %>% 
  mutate(measure = as.factor(measure),
         t = as.integer(t))

# Absolute inputs
N_inputs_df <- N_inputs_df %>% 
  mutate(mean = mean %>% set_units("g") %>% set_units("gN"),
         sd = sd %>% set_units("g") %>% set_units("gN"))

# Inputs relative to biomass produced
N_inputs_rel_df <- N_inputs_df %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(mean = mean.x, 
         sd = sd.x, 
         biomass_mean = mean.y,
         biomass_sd = sd.y) %>% 
  mutate(biomass_mean = biomass_mean %>% set_units("g_fish"),
         biomass_sd = biomass_sd %>% set_units("g_fish"),
         biomass_sd = biomass_sd/biomass_mean,
         sd = sd/mean,
         mean = mean/biomass_mean,
         sd = sqrt(sd^2 + biomass_sd^2) * mean,
         biomass_sd = biomass_sd * biomass_mean)
```

```{r fig-composition-comparison}
#| eval: false
#| fig-cap: Comparison of the composition of the reference (blue), past (red) and future (green) feed and faeces of salmon fed each feed.

ggplot(ec, aes(x = source, y = 100*value, fill = feed)) +
  geom_col(colour = "black", position = position_dodge(width = 0.9), alpha = 0.75) +
  facet_grid(cols = vars(macro)) +
  prettyplot() +
  scale_fill_manual(values = feed_pal) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  labs(y = "Composition (%)", x = "Source")
```

```{r composition stats}
# Through time
N_inputs_ts <- N_inputs_df %>%
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

N_inputs_rel_ts <- N_inputs_rel_df %>% 
  group_by(feed, t, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

# Total (at production end)
N_inputs_end <- N_inputs_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g"),
          mean = meanna(mean))

N_inputs_rel_end <- N_inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(mean),
          max = maxna(mean),
          sd = sdna(mean) %>% set_units("g g_fish-1"),
          mean = meanna(mean))

N_inputs_rel_end_diff <- N_inputs_rel_df %>% 
  group_by(feed, measure, farm_ID) %>% 
  reframe(mean = sumna(mean)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(future = (future-reference)/reference,
         past = (past-reference)/reference) %>% 
  select(-reference) %>% 
  pivot_longer(cols = c(past, future), names_to = "feed", values_to = "diff") %>% 
  mutate(feed = factor(feed, levels = c("past", "future"))) %>% 
  group_by(feed, measure) %>% 
  reframe(min = minna(diff),
          max = maxna(diff),
          sd = sdna(diff),
          mean = meanna(diff))

# Pull key values (uneaten feed)
u_stmin1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(min)
u_stmax1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(max)
u_stmean1 <- filter(N_inputs_rel_end, feed == "reference", measure == "uneat") %>% pull(mean)
u_stmin2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(min)
u_stmax2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(max)
u_stmean2 <- filter(N_inputs_rel_end_diff, feed == "past", measure == "uneat") %>% pull(mean)
u_stmin3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(min)
u_stmax3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(max)
u_stmean3 <- filter(N_inputs_rel_end_diff, feed == "future", measure == "uneat") %>% pull(mean)

# Convert past and future values to % change values
u_stmin2 <- (100*u_stmin2) %>% round(2)
u_stmax2 <- (100*u_stmax2) %>% round(2)
u_stmean2 <- (100*u_stmean2) %>% round(2)
u_stmin3 <- (100*u_stmin3) %>% round(2)
u_stmax3 <- (100*u_stmax3) %>% round(2)
u_stmean3 <- (100*u_stmean3) %>% round(2)

u_stmin1 <- u_stmin1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmax1 <- u_stmax1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
u_stmean1 <- u_stmean1 %>% set_units("g kg_fish-1") %>% drop_units() %>% round(2)
```

# Supplementary materials

## Parameter sensitivity {#sec-parameter-sensitivity}

```{r sensitivity data}
sens_weight <- output_sens_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("results") %>% 
  str_subset("weight") %>% 
  qs::qread()

sens_excr <- output_sens_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("results") %>% 
  str_subset("total_excr") %>% 
  qs::qread()

sens_data <- rbind(sens_weight, sens_excr)
```

The most sensitive parameter by far in @fig-param-sens-weight is $m$ (the weight exponent for anabolism), followed closely by $n$ the (weight exponent for catabolism). 
These two are also relatively uncertain, as $m$ was taken from *Salmo trutta* measurements in a fairly old paper [@broekhuizen_modelling_1994] and $n$ was taken from @baldan_r_2018 which did not specifically look at Atlantic salmon.

```{r fig-param-sens-weight}
#| fig-cap: Impact of increasing each model parameter by 10% on final farm biomass (red) and total excreted nutrients (blue) over the whole production period.

sens_data <- sens_data %>% 
  filter(measure %in% c("total_excr", "total_uneat"))

sens_data %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens - sd_sens, ymax = mean_sens + sd_sens, fill = measure)) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black", alpha = 0.75) +
  # geom_errorbar(position = position_dodge(width = 0.95), width = 0.3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_y_continuous(breaks = seq(-30,30,5), limits = c(-10,30)) +
  # scale_x_discrete(limits = rev(param_names$name), labels = rev(param_names$lab)) +
  # scale_fill_manual(values = c(total_excr = "salmon", total_uneat = "steelblue")) +
  labs(x = "Adjusted parameter", y = "Sensitivity") +
  prettyplot() +
  theme(strip.text = element_blank(), 
        aspect.ratio = 0.8,
        axis.title.y = element_blank()) +
  coord_flip()
```

# References