### 2025-08-15 index.qmd
Small QOL improvements

---
title: "Local to Global Mariculture Modelling"
subtitle: "Project Documentation"
author: "Dr Tormey Reimer"
date: today
execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true # quarto render index.qmd --cache-refresh
knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 8.5
format:
  html:
    code-fold: true
---

# Table of Contents

All code and processes for this project are linked below. Note that some scripts require access to raw data which is not stored on Github. 
While this page (the index) has code that runs prior to rendering, all code in the subsequent pages has not been run for rendering.

## Main scripts

The main scripts for this project focus on setting up the model to examine the differences in total protein, carbohydrate, lipid and nitrogen losses between feeds.

| Page | Description |
|---------|:-------------------------------------------------|
| **[Data Tidying](01_tidying_data.html)** | Initial data tidying and preparation |
| **[Extracting Production Locations](02_extracting_production_locations.html)** | Identification and processing of mariculture production site locations |
| **[Assessing Species Potential](03_assessing_species_potential.html)** | Evaluation of species suitability for different locations |
| **[Extracting Temperatures](04_extracting_temperatures.html)** | SST data extraction and processing for model inputs |
| **[Formulating Feeds](05_formulating_feeds.html)** | Feed composition formulation for model runs |
| **[Running Atlantic Salmon Model](06_run_atlantic_salmon_model.html)** | Atlantic salmon growth model execution |

## Extra scripts

The extra scripts for this project involved some exploration into potential impacts of nutrient losses. These will likely not make it into the final manuscript.

| Page | Description |
|---------|:-------------------------------------------------|
| **[Species Layers](07_species_layers.html)** | Spatial analysis of species distribution layers, to try and acertain impacts on vulnerable species |
| **[Relative N Inputs](08_relative_N_inputs.html)** | Analysis of nitrogen input (runoff) patterns and the relative impacts of farm nutrients |

```{r setup}
#| cache: false
#| message: false

library(magrittr)
library(units)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(purrr)
library(future)
library(readxl)
library(conflicted)
library(lubridate)
library(RColorBrewer)
# library(lme4)
library(car)
library(emmeans)

conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect())
# This markdown uses TinyTex - install with tinytex::install_tinytex()

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% walk(source)

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate", "worrms", "aquamapsdata") %>% sort()
```

```{r}
#| label: global-variables
#| include: false
#| cache: false

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")

species_params <- find_read(output_species_data_path, "species_params")
pop_params <- find_read(output_species_data_path, "pop_params")

farm_locs <- find_read(output_farm_data_path, "geometry") %>% 
  rename(farm_ID = farm_id) %>% 
  mutate(country = as.factor(country))
countries_long <- farm_locs %>%
  mutate(
    longitude = map_dbl(geometry, ~st_coordinates(.x)[1]),
    latitude = map_dbl(geometry, ~st_coordinates(.x)[2])
  ) %>% 
  st_drop_geometry() %>% 
  group_by(country) %>% 
  reframe(latitude = mean(latitude)) %>% 
  arrange(-latitude) %>% 
  pull(country) %>% 
  as.character() 
countries_short <- countries_long %>% 
  str_replace("United States of America", "USA") %>% 
  str_replace("Russian Federation", "Russia") %>% 
  str_replace("United States of America", "USA")
farm_locs <- farm_locs %>% 
  mutate(
    country = factor(
      country, 
      levels = countries_long,
      labels = countries_short
    ))

long_feeds <- c("marine_dominant", "plant_dominant", "novel_inclusive")
short_feeds <- c("MD", "PD", "NI")

feed_pal <- c(
  "past" = "#E41A1C", 
  "reference" = "#377EB8", 
  "future" = "#4DAF4A",
  "plant_dominant" = "#4DAF4A",
  "marine_dominant" = "#377EB8",
  "novel_inclusive" = "#E41A1C",
  "PD" = "#4DAF4A",
  "MD" = "#377EB8",
  "NI" = "#E41A1C"
)
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", 
  "P" = "darkred", 
  "carbohydrates" = "steelblue", 
  "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", 
  "L" = "darkgoldenrod1"
)

prettyplot <- function() {
  theme_classic() +
    theme(
      text = element_text(family = "serif", size = 14, colour = "black"),
      legend.position = "none",
      axis.title.y = element_text(margin = margin(r = 5)),
      axis.title.x = element_text(margin = margin(t = 5))
    )
}
```

# Results {#sec-results}

```{r}
#| label: load-all-data-analysis

# Biomass produced
bpr1 <- find_read(output_model_farm_path, "biomass_produced") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(mean = mean(mean) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = mean)

bpr2 <- find_read(comparisons_path, "biomass_produced") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(mean)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Feed provided
fpr1 <- find_read(data_analysis_path, "food_prov_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

fpr2 <- find_read(comparisons_path, "food_prov_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Feed provided per biomass
fpr_pb1 <- find_read(data_analysis_path, "food_prov_farm_per_biom_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(mean_biom = mean(mean_biom) %>% set_units("g g_fish-1") %>% set_units("kg kg_fish-1") %>% drop_units() %>% round(2)) %>% 
  pivot_wider(names_from = feed, values_from = mean_biom)

fpr_pb2 <- find_read(comparisons_path, "food_prov_farm_per_biom_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(mean_biom)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Uneaten feed
tuf1 <- find_read(data_analysis_path, "total_uneat_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tuf2 <- find_read(comparisons_path, "total_uneat_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Excreted waste
tew1 <- find_read(data_analysis_path, "total_excr_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tew2 <- find_read(comparisons_path, "total_excr_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Total waste
tow1 <- find_read(data_analysis_path, "total_mass_inputs_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tow2 <- find_read(comparisons_path, "total_mass_inputs_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Excretion proportion of waste
epw <- find_read(comparisons_path, "excr_proportion_of_waste") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Protein, lipid, carbs
plc1 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed, measure, macro) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro), values_from = total)

# Proportion of PLC waste from excretion
plc2 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed, measure, macro) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro, measure), values_from = total) %>% 
  mutate(
    MD_C = round(100*MD_C_excr/(MD_C_excr+MD_C_uneat),1),
    MD_L = round(100*MD_L_excr/(MD_L_excr+MD_L_uneat),1),
    MD_P = round(100*MD_P_excr/(MD_P_excr+MD_P_uneat),1),
    NI_C = round(100*NI_C_excr/(NI_C_excr+NI_C_uneat),1),
    NI_L = round(100*NI_L_excr/(NI_L_excr+NI_L_uneat),1),
    NI_P = round(100*NI_P_excr/(NI_P_excr+NI_P_uneat),1),
    PD_C = round(100*PD_C_excr/(PD_C_excr+PD_C_uneat),1),
    PD_L = round(100*PD_L_excr/(PD_L_excr+PD_L_uneat),1),
    PD_P = round(100*PD_P_excr/(PD_P_excr+PD_P_uneat),1)
  ) %>% 
  select(-contains(c("excr", "uneat")))

# % change in PLC waste
plc3 <- find_read(comparisons_path, "PLC_uneat_excr_farm_summed") %>% 
  group_by(feed, macro) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro), values_from = diff)

# N inputs, uneaten
niu1 <- find_read(data_analysis_path, "P_uneat_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds),
         total = total * 6.25,
         sd = sd * 6.25) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

niu2 <- find_read(comparisons_path, "N_uneat_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# N inputs, excreted
nie1 <- find_read(data_analysis_path, "P_excr_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds),
         total = total * 6.25,
         sd = sd * 6.25) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

nie2 <- find_read(comparisons_path, "N_excr_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# N inputs, total
nit1 <- find_read(data_analysis_path, "N_total_farm_summed.qs") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

nit2 <- find_read(comparisons_path, "N_total_farm_summed.qs") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)
```

@tbl-results-summary is a summary of all the results I've processed so far, basically trying to acertain all the major differences in inputs between the feeds.

The NI diet was more efficient than the PD diet. When farms were fed the NI diet they had an average FCE* of `r fpr_pb1$NI`, compared to the average FCE of `r fpr_pb1$PD` when farms were fed the PD diet, and feeding the NI diet resulted in `r bpr2$NI`% more biomass produced with the same population compared to the PD diet.

::: {.aside}
*Note that when I refer to "FCE" here I'm talking about the "PLC weight" (weight of protein + lipid + carbohydrate) of feed, which may differ from its actual weight because of the inclusion of minerals and non-digestible material.
:::

Despite this, nutrients lost to the environment were generally highest in the NI diet and lowest in the MD diet. 
In terms of total waste, the NI and PD diets were very similar. I initially thought that this was because the NI and PD diets are more similar than the MD and PD diets, but this doesn't match up with the numbers (PD and NI diets are 63.3% the same, while the PD and MD diets are 54.3% the same). 
The MD diet had markedly lower total waste than either the PD or NI diets.

However, the composition of waste lost to the environment was very different between the diets, particularly in terms of proteins and carbohydrates (@fig-PLC-waste-comparison). The PD diet had the lowest protein waste output and the NI diet had the highest, while the MD diet had the lowest carbohydrate and lipid waste outputs and the PD diet the highest. Overall, changing from the PD diet to the NI diet would result in a `r abs(plc3$NI_P)`% increase in protein waste but a `r abs(plc3$NI_C)`% and `r abs(plc3$NI_L)`% reduction in carbohydrates and lipids respectively, while changing to the MD diet would result in a `r abs(plc3$MD_P)`% increase in protein waste but a `r abs(plc3$NI_C)`% and `r abs(plc3$NI_L)`% reduction in carbohydrates and lipids respectively.

Consequently, while total waste was only `r tow2$NI`% higher when farms were fed the NI diet, protein waste was `r plc3$NI_P`% higher and nitrogen waste was `r nit2$NI`% higher relative to when farms were fed the PD diet.

Proteins in the NI diet were only slightly less digestible than those used in the PD diet (@tbl-modelled-feeds), but this difference, combined with the higher protein content in the NI diet, resulted in the large difference in nitrogen lost to the environment.

| Description | Plant-dominant | Marine-dominant | Novel-inclusive | Change from PD to MD (%) | Change from PD to NI (%) |
|:--------------------------|:----------:|:----------:|:----------:|:----------:|:----------:|
| Salmon biomass produced (t)                         | `r bpr1$PD`        | `r bpr1$MD`        | `r bpr1$NI`        | `r coln(bpr2$MD)`    | `r coln(bpr2$NI)`    |
| Feed provided (t)                                   | `r fpr1$PD`        | `r fpr1$MD`        | `r fpr1$NI`        | `r coln(fpr2$MD)`    | `r coln(fpr2$NI)`    |
| Feed provided (kg/kg salmon)                        | `r fpr_pb1$PD`     | `r fpr_pb1$MD`     | `r fpr_pb1$NI`     | `r coln(fpr_pb2$MD)` | `r coln(fpr_pb2$NI)` |
| Uneaten feed (t)                                    | `r tuf1$PD`        | `r tuf1$MD`        | `r tuf1$NI`        | `r coln(tuf2$MD)`    | `r coln(tuf2$NI)`    |
| Excreted waste (t)                                  | `r tew1$PD`        | `r tew1$MD`        | `r tew1$NI`        | `r coln(tew2$MD)`    | `r coln(tew2$NI)`    |
| Total waste (E+U, t)                                | `r tow1$PD`        | `r tow1$MD`        | `r tow1$NI`        | `r coln(tow2$MD)`    | `r coln(tow2$NI)`    |
| Proportion of total waste composed of excretion (%) | `r epw$PD`         | `r epw$MD`         | `r epw$NI`         |                      |                      |
| Total protein waste (t)                             | `r sum(plc1$PD_P)` | `r sum(plc1$MD_P)` | `r sum(plc1$NI_P)` | `r coln(plc3$MD_P)`  | `r coln(plc3$NI_P)`  |
| Total carbohydrate waste (t)                        | `r sum(plc1$PD_C)` | `r sum(plc1$MD_C)` | `r sum(plc1$NI_C)` | `r coln(plc3$MD_C)`  | `r coln(plc3$NI_C)`  |
| Total lipid waste (t)                               | `r sum(plc1$PD_L)` | `r sum(plc1$MD_L)` | `r sum(plc1$NI_L)` | `r coln(plc3$MD_L)`  | `r coln(plc3$NI_L)`  |
| Proportion of protein waste from excretion (%)      | `r plc2$PD_P`      | `r plc2$MD_P`      | `r plc2$NI_P`      |                      |                      |
| Proportion of carbohydrate waste from excretion (%) | `r plc2$PD_C`      | `r plc2$MD_C`      | `r plc2$NI_C`      |                      |                      |
| Proportion of lipid waste from excretion (%)        | `r plc2$PD_L`      | `r plc2$MD_L`      | `r plc2$NI_L`      |                      |                      |
| Total N from uneaten feed (t)                       | `r niu1$PD`        | `r niu1$MD`        | `r niu1$NI`        | `r coln(niu2$MD)`    | `r coln(niu2$NI)`    |
| Total N from excreted waste (t)                     | `r nie1$PD`        | `r nie1$MD`        | `r nie1$NI`        | `r coln(nie2$MD)`    | `r coln(nie2$NI)`    |
| Total N inputs (E+U, t)                             | `r nit1$PD`        | `r nit1$MD`        | `r nit1$NI`        | `r coln(nit2$MD)`    | `r coln(nit2$NI)`    |
: Summary of all results. Unless otherwise specified, values presented are means across all farms, summed across a single production period. In the % change columns, the colour of the text indicates whether there is an increase (red) or a decrease (green). Some cells of this table are blank - sometimes that means it doesn't make sense to fill that cell, but sometimes it means I haven't got to it yet. {#tbl-results-summary}

```{r}
#| label: PLC-waste-comparison

plc1 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed")

lm_P <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "P"))
aov_P <- aov(lm_P)
# summary(aov_P)

emm <- emmeans(lm_P, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)

lm_C <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "C"))
aov_C <- aov(lm_C)
# summary(aov_C)

emm <- emmeans(lm_C, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)

lm_L <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "L"))
aov_L <- aov(lm_L)
# summary(aov_L)

emm <- emmeans(lm_L, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)
```

::: {.aside}
I accidentally used raw total waste in @fig-PLC-waste-comparison instead of waste per kg of biomass
:::

```{r}
#| label: fig-PLC-waste-comparison
#| fig-cap: Mean waste lost to the environment in a complete production cycle when farms are fed the MD (blue), PD (green) or NI (red) diet. Errorbars show SD across all farms.

plc1 %>%
  group_by(measure, macro, feed) %>%
  reframe(
    sd = sd(total) %>% set_units("g") %>% set_units("t") %>% drop_units(),
    total = mean(total) %>% set_units("t") %>% drop_units(), 
  ) %>% 
  mutate(
    macro = factor(
      macro, 
      levels = c("P", "C", "L"),
      labels = c("Protein", "Carbohydrate", "Lipid")
    ),
    measure = factor(
      measure,
      levels = c("uneat", "excr"),
      labels = c("Uneaten", "Excreted")
    )
  ) %>% 
  ggplot(aes(
    x = measure, 
    y = total, 
    ymin = total-sd, ymax = total+sd,
    fill = feed
  )) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  facet_grid(cols = vars(macro)) +
  prettyplot() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Mean total waste (t)", x = "Source")
```

# N inputs only

```{r}
#| label: N-inputs-comparisons

N_uneat <- find_read(comparisons_path, "N_uneat_farm_per_biom_summed")
N_excr <- find_read(comparisons_path, "N_excr_farm_per_biom_summed")

N_all <- rbind(N_uneat, N_excr) %>% 
  st_drop_geometry() %>% 
  mutate(
    measure = droplevels(measure),
    country = factor(country,
                   levels = countries_long,
                   labels = countries_short)
  )

# ggplot(N_all, aes(x = country, y = mean_biom, fill = feed)) +
#   geom_boxplot() +
#   facet_grid(cols = vars(measure)) +
#   scale_fill_manual(values = feed_pal)
```

# Spatially

```{r}
#| label: fig-spatial-N-inputs
#| fig-height: 12
#| fig-cap: Total N inputs per unit biomass of salmon produced when farms are fed the MD feed (top), the PD feed (middle) or the NI feed (bottom).

N_ins <- rbind(
  find_read(data_analysis_path, "P_excr_farm_per_biom_summed"),
  find_read(data_analysis_path, "P_uneat_farm_per_biom_summed")
) %>% 
  st_drop_geometry() %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(country, farm_ID, feed) %>% 
  reframe(
    mean_biom = sum(mean_biom) %>% set_units("kg t_fish-1") %>% drop_units(), 
    sd_biom = sqrt(sum(sd_biom^2)) %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  merge(farm_locs, by = c("country", "farm_ID")) %>% 
  st_as_sf()

p_bigmap_robinson +
  geom_sf(
    data = N_ins, 
    aes(colour = mean_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(0, 120, 10)) + 
  guides(col = guide_colourbar(
    title = expression("Total N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  facet_grid(rows = vars(feed)) +
  # prettyplot() +
  theme(strip.background = element_blank(), strip.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "bottom")
```

```{r}
#| label: N-inputs-patchwork-map

p_mercator <- p_bigmap_mercator +
  geom_sf(
    data = N_ins,
    aes(colour = mean_biom, geometry = geometry), 
    size = 2.5
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(0, 120, 10)) + #, limits = c(-30, 10)
  guides(col = guide_colourbar(
    title = expression("Total N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  # facet_wrap(facets = vars(feed), nrow = 2, ncol = 2) +
  prettyplot() +
  theme(
    # strip.background = element_blank(), strip.text = element_blank(),
    legend.position = "bottom"
  ) +
  labs(y = "Latitude", x = "Longitude")

ins <- get_insets_mercator(p_mercator)

ins[[1]] + facet_grid(cols = vars(feed))
ins[[2]] + facet_grid(rows = vars(feed))
ins[[3]] + facet_grid(cols = vars(feed))
ins[[4]] + facet_grid(cols = vars(feed))
ins[[5]] + facet_grid(cols = vars(feed))
```

# Supplemental methods {#sec-supp-methods}

```{r}
#| label: location-data

farm_locs <- find_read(output_farm_data_path, "geometry") %>% 
  rename(farm_ID = farm_id)

farm_mean_lat <- farm_locs %>%
  st_centroid() %>%
  mutate(latitude = st_coordinates(.)[, "Y"]) %>%
  st_drop_geometry() %>%
  group_by(country) %>%
  summarise(mean_latitude = mean(latitude, na.rm = TRUE)) %>% 
  arrange(-mean_latitude)

countries <- farm_mean_lat$country
countries[str_detect(countries, "United States of America")] <- "USA"
countries[str_detect(countries, "United Kingdom")] <- "UK"
countries[str_detect(countries, "Russian Federation")] <- "Russia"

farm_locs <- farm_locs %>% 
  mutate(country = factor(country, levels = farm_mean_lat$country, labels = countries))
```

## Population and biomass {#sec-population-biomass}

```{r}
#| label: raw biomass and weight data

weight_cohort <- find_read(output_model_cohort_path, "weight_stat_all_farms")
biomass_cohort <- find_read(output_model_cohort_path, "biomass_stat_all_farms")
```

Each farm's final population was calculated by dividing the production tonnage required by the average modelled individual harvest size for each farm (when feed reference feed, @fig-harvest-size). A population timeseries was then generated using the standard mortality rate (`r round(unname(pop_params['mortmyt'])*100, 3)` % d$^{-1}$).

```{r}
#| label: fig-harvest-size
#| fig-cap: |
#|   Frequency distribution of harvest sizes used to create the population timeseries for each farm, as predicted by the fish-growth model.

harv_size <- find_read(output_farm_data_path, "farm_harvest_size")

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg") %>% drop_units()) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  scale_x_continuous(breaks = seq(0,7,0.5)) +
  labs(y = "Frequency", x = "Harvest weight (kg)") +
  prettyplot() +
  theme(aspect.ratio = 0.75)
```

Using this method, a farm's 'population' is not a good representation of the actual number of individuals being farmed, but it ensures each farm location is producing the correct tonnage while allowing comparisons between the different feeds.

```{r}
#| label: fig-cohorts
#| fig-cap: | 
#|   Example of how the overlapping cohorts combine into a single farm production curve (in this case biomass). The left plot shows the farm biomass produced by each cohort from initialisation to harvest, while the right plot shows the combined farm biomass over the course of a single production period of 548 days.

farm_biomass <- find_read(output_model_farm_path, "biomass_stat") %>% 
  filter(farm_ID == 722 & feed == "plant_dominant")

cohort_biomass_split <- rbind(
  farm_biomass %>% 
    farm_to_cohort() %>% 
    mutate(cohort = 1),
  farm_biomass %>% 
    farm_to_cohort(time_offset = 365) %>% 
    mutate(cohort = 2),
  farm_biomass %>% 
    farm_to_cohort(time_offset = 730) %>% 
    mutate(cohort = 3)
) %>% 
  mutate(split = 0) %>% 
  relocate(farm_ID, .before = t) %>% 
  relocate(feed, .before = t)

cohort_biomass <- find_read(output_model_cohort_path, "biomass_stat") %>% 
  filter(farm_ID == 722 & feed == "plant_dominant") %>% 
  mutate(cohort = "all", split = 1)

rbind(
  cohort_biomass_split,
  cohort_biomass
  ) %>% 
  mutate(
    t = t-min(t)+1,
    mean = mean %>% set_units("g") %>% set_units("t"),
    sd = sd %>% set_units("g") %>% set_units("t")
  ) %>% 
  ggplot(aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  facet_grid(cols = vars(split)) +
  scale_x_continuous(breaks = seq(0, 1600, 120)) +
  scale_colour_manual(values = cohort_pal) +
  prettyplot() +
  theme(strip.background = element_blank(), strip.text = element_blank()) +
  labs(y = "Farm biomass", x = "Day of production")
```

```{r}
#| label: pop-biom-data

biomass_farm <- find_read(output_model_farm_path, "biomass_stat") %>% 
  group_by(farm_ID, feed) %>% 
  slice_max(t) %>% 
  ungroup()

production <- find_read(output_farm_data_path, "farm_static_data") %>% 
  merge(biomass_farm, by = "farm_ID", all.x = F) %>% 
  mutate(
    mean = set_units(mean, "g"),
    diff = set_units(mean - tonnes_per_farm, "t"),
    perc_diff = diff/tonnes_per_farm
  ) %>% 
  group_by(country, feed) %>% 
  reframe(
    min_percdiff = min(perc_diff),
    max_percdiff = max(perc_diff),
    mean_percdiff = mean(perc_diff)
  )
```

By this method, farm biomass as predicted by the model predicted farm tonnage relatively accurately. Predicted farm biomass was within `r round(100*min(production$min_percdiff),1)`% -- `r round(100*max(production$max_percdiff),1)`% of tonnage produced.

## Feed provision and feeding response to temperature {#sec-model-feeding}

@fig-functional-response-to-temperature shows the response of food ingested to temperature, and the food provided in response.

```{r}
#| label: fig-functional-response-to-temperature
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dotted line), relative to maximum potential ingestion. The vertical dotted lines show the actual minimum and maximum daily temperatures experienced by farms represented in the model.

Sys.setenv(TAR_PROJECT = "project_farmruns")
farm_ts_data <- tar_read(farm_ts_data_chunked) %>% bind_rows()
min_T <- min(farm_ts_data$temp_c)
max_T <- max(farm_ts_data$temp_c)

all_params <- c(
  find_read(output_species_data_path, "species_params"),
  find_read(output_species_data_path, "pop_params")
)

func_T_resp <- data.frame(water_temp = seq(-5, 26, 0.25))
weight <- 1000

# Relative feeding rate with temperature
func_T_resp$feeding_rate <- sapply(
  FUN = feeding_rate, 
  X = func_T_resp$water_temp, 
  species_params = all_params
)

# Maximum potential ingestion (g) at temperature
func_T_resp$ing_pot <- all_params['meanImax'] * (weight^all_params['m']) * func_T_resp$feeding_rate

# Food provided is based on calculated potential ingestion BUT if temperature is too low they are still provided with the food that would be required at lowest temperature (when feeding rate is ~18.3%)
func_T_resp$food_prov <- food_prov_rate(
  water_temp = func_T_resp$water_temp, 
  ing_pot = func_T_resp$ing_pot,
  ing_pot_min = all_params['meanImax'] * (weight^all_params['m']) * feeding_rate(all_params['Taa'], all_params),
  species_params = all_params,
  pop_params = all_params
)

func_T_resp$food_enc <- all_params['eff'] * func_T_resp$food_prov
func_T_resp <- func_T_resp %>% 
  mutate(ing_act = pmin(food_enc, ing_pot))

max_ing <- max(func_T_resp$ing_act)

func_T_resp %>% 
  ggplot(aes(x = water_temp, y = ing_act/max_ing)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov/max_ing), linewidth = 0.75, linetype = "dotted") +
  geom_vline(xintercept = min_T, linetype = "dotted") +
  geom_vline(xintercept = max_T, linetype = "dotted") +
  # scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Feedd ingested/provided") +
  prettyplot()
```

```{r}
#| label: patchwork-map-mercator
#| eval: false
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

p_bigmap_feeding <- p_bigmap_mercator +
  geom_sf(data = feeding, aes(color = prop_feed_eaten, geometry = geometry), size = 2) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Proportion of provided feed eaten",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

patchwork_mercator(p_bigmap_feeding)
```

## Modelled experimental feeds {#sec-feeds}

```{r}
#| label: feed-details

feed_params <- find_read(output_species_data_path, "feed_params")

feed_summary <- map_dfr(names(feed_params), function(feed) {
  df <- feed_params[[feed]]
  map_dfr(c("Proteins", "Lipids", "Carbohydrates"), function(macro) {
    contrib <- df[[macro]]$proportion * df[[macro]]$macro
    total <- sum(contrib)
    digest <- sum(df[[macro]]$digest * contrib) / total
        tibble(
      feed = feed,
      macro = macro,
      total = total,
      digest = digest
    )
  })
}) %>%
  mutate(macro = factor(macro, levels = c("Proteins", "Lipids", "Carbohydrates"), labels = c("P", "L", "C")),
         feed = as.factor(feed),
         total_r = total * 100 %>% round(1),
         digest_r = digest * 100 %>% round(1))
```

| Feed | Protein | Lipid | Carbohydrate |
|------|:----------:|:----------:|:----------:|
| Plant-dominant | `r round(100*filter(feed_summary, feed == "plant_dominant", macro == "P")$total, 1)`% (`r round(100*filter(feed_summary, feed == "plant_dominant", macro == "P")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "plant_dominant", macro == "L")$total, 1)`% (`r round(100*filter(feed_summary, feed == "plant_dominant", macro == "L")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "plant_dominant", macro == "C")$total, 1)`% (`r round(100*filter(feed_summary, feed == "plant_dominant", macro == "C")$digest, 1)`%) |
| Marine-dominant | `r round(100*filter(feed_summary, feed == "marine_dominant", macro == "P")$total, 1)`% (`r round(100*filter(feed_summary, feed == "marine_dominant", macro == "P")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "marine_dominant", macro == "L")$total, 1)`% (`r round(100*filter(feed_summary, feed == "marine_dominant", macro == "L")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "marine_dominant", macro == "C")$total, 1)`% (`r round(100*filter(feed_summary, feed == "marine_dominant", macro == "C")$digest, 1)`%) |
| Novel-inclusive | `r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "P")$total, 1)`% (`r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "P")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "L")$total, 1)`% (`r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "L")$digest, 1)`%) | `r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "C")$total, 1)`% (`r round(100*filter(feed_summary, feed == "novel_inclusive", macro == "C")$digest, 1)`%) |
: Feed macronutrient details (composition in feed, digestibility). {#tbl-modelled-feeds}

## Model sensitivity to parameter changes {#sec-param-sensitivity}

Parameter sensitivity was tested by varying each parameter by $\pm 10$% and calculating sensitivity as $S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}$, where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively. 
The population-level parameters ($\overline{W}$, $\Delta{W}$, $\overline{I_{max}}$, $\Delta{I_{max}}$, $\overline{F}$, $\Delta{F}$) were tested with a population of of 5000 on a random 10% of farms, while all other parameters were tested across all farms with a population of 10 (@fig-sensitivity-results).

```{r}
#| label: salmon-param-names

params <- find_read(output_species_data_path, "sens_params")

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

```{r}
#| label: fig-sensitivity-results
#| fig-cap: Sensitivity of modelled individual fish weight (blue) and total farm excretion across a full production period (red) to a 10% increase in each model parameter. Bars show mean values, error bars show standard deviation between farms. 
#| fig-width: 8

sens_results <- find_read(output_sens_data_path, "sens_results") %>% 
  filter(measure %in% c("total_excr", "weight") & adj_param != "nruns") %>% 
  mutate(measure = droplevels(measure))

sens_results %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens, fill = measure)) +
  geom_col(position = position_dodge(), width = 0.95, alpha = 0.65, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  prettyplot() +
  scale_x_discrete(breaks = param_names$name, labels = param_names$lab) +
  scale_y_continuous(breaks = seq(-10, 30, 5)) +
  scale_fill_brewer(palette = "Set1") +
  labs(y = "Sensitivity", x = "Adjusted parameter")
```

# Code and packages used {#sec-code}

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `r cite_packages(packages)`.

For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].