### 2025-09-04 archive/04_atlantic_salmon_synthesis/04_atlantic_salmon_synthesis.qmd

---
title: "Synthesising the model outputs"
format: html
editor: source

execute:
  eval: false
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup}
#| code-summary: Setup

# library(tidyverse)
library(dplyr)
library(arrow)
library(magrittr)
library(here)
library(qs)
library(sf)
library(dtplyr)
library(future)
library(furrr)
library(stringr)
library(conflicted)
conflicts_prefer(dplyr::filter(), dplyr::select())

this_species <- "atlantic_salmon"
```

```{r locations files}
#| code-summary: Bring in spatial file with farm IDs

# locations_file <- qread(file = sprintf("data/_general_data/farm_locations/%s_locations_w_temps.qs", this_species)) %>% 
#   select(-c(day, temp_c)) %>% 
#   group_by(farm_id) %>% 
#   group_split() %>% 
#   map_df(.f = \(x){return(x %>% slice(1))})

locations_file <- qread(file = sprintf("data/_general_data/farm_locations/%s_locations_w_temps.qs", this_species)) %>% 
  select(-c(day, temp_c)) %>% 
  group_by(farm_id) %>% 
  slice_head(n = 1) %>% 
  ungroup()
```

We are primarily interested in biomass produced, excretion (ammonia), faeces, and wasted feed broken down to nitrogen.

REFERENCE FEED

Bring in relevant model outputs.

```{r}
this_feed <- "reference"

biomass_files <- list.files(path = here("data/atlantic_salmon/data_products/model_outputs/"), pattern = "biomass_", full.names = TRUE)

faeces_files <- list.files(path = here("data/atlantic_salmon/data_products/model_outputs/"), pattern = "excretion_", full.names = TRUE)

feed_waste_files <- list.files(path = here("data/atlantic_salmon/data_products/model_outputs/"), pattern = "feed_waste_", full.names = TRUE)

ammonia_files <- list.files(path = here("data/atlantic_salmon/data_products/model_outputs/"), pattern = "NH4_", full.names = TRUE)

weight_files <- list.files(path = here("data/atlantic_salmon/data_products/model_outputs/"), pattern = "weight_", full.names = TRUE)
```

Reference feeds - loop through the locations file to add the biomass, excretion, faeces, uneaten feed in N.

```{r}
feed_type <-  "reference"

commercial_size <- read.csv(sprintf(here("data/%s/params/Parameters_%s.csv"), this_species, feed_type)) %>% 
  filter(Description == "Commercial size") %>% 
  pull(Value) %>% 
  as.numeric()

farm_ids <- locations_file$farm_id
# this_id <- farm_ids[[29]]

#Set up multicore
future::plan(strategy = "multisession", workers = parallel::detectCores()-2)

farm_N_reference <- future_map(.x = farm_ids, .f = \(this_id){
  message("Synthesising nitrogen data at harvest this farmID_", this_id)

  weight_days <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(days)
  weights <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(weight)

  harvest_day <- 
    weight_days %>% 
    pluck(if(tail(weights, n=2)[1]>commercial_size){
      head(which(weights>commercial_size), n =2)[1]
    } else {
      tail(weight_days, n=2)[1]
      } )

# biomass
this_biomass_mean <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
  slice(harvest_day) %>% pull(biomass)/1e+6

this_biomass_lb <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
  slice(harvest_day) %>% pull(lower_bound)/1e+6

this_biomass_ub <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
  slice(harvest_day) %>% pull(upper_bound)/1e+6

# Weight
this_weight_mean <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(weight)

this_weight_lb <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(lower_bound)

this_weight_ub <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day)%>% pull(upper_bound)

# faeces 
this_faeces_mean <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_excretion) %>% sum()/6.25/1e+6

this_faeces_lb <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

this_faeces_ub <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

# feed waste
this_feed_waste_mean <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_waste) %>% sum()/6.25/1e+6

this_feed_waste_lb <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

this_feed_waste_ub <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

# excretion (NH4)
this_ammonia_mean <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
  slice(1:harvest_day) %>% pull(nh4_production) %>% sum()/1e+6

this_ammonia_lb <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
  slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/1e+6

this_ammonia_ub <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
  slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/1e+6

return(data.frame(
  farm_id = this_id, 
  biomass_mean = this_biomass_mean, 
  biomass_lb = this_biomass_lb, 
  biomass_ub = this_biomass_ub, 
  faeces_mean = this_faeces_mean, 
  faeces_lb = this_faeces_lb, 
  faeces_ub = this_faeces_ub, 
  excretion_mean = this_ammonia_mean, 
  excretion_lb = this_ammonia_lb, 
  excretion_ub = this_ammonia_ub, 
  feed_waste_mean = this_feed_waste_mean, 
  feed_waste_lb = this_feed_waste_lb, 
  feed_waste_ub = this_feed_waste_ub, 
  total_N_mean = sum(this_faeces_mean, this_ammonia_mean, this_feed_waste_mean), 
  total_N_lb = sum(this_faeces_lb, this_ammonia_lb, this_feed_waste_lb), 
  total_N_ub = this_feed_waste_ub
))
}) %>% bind_rows()

farm_N_reference_spatial <- 
  locations_file %>% 
  select(iso3c, country, tonnes_per_farm, geometry, farm_id) %>% 
  distinct() %>% 
  left_join(farm_N_reference)

qsave(x = farm_N_reference_spatial, file =  sprintf(here("data/atlantic_salmon/data_products/model_summaries/%s/nitrogen_summary.qs"), this_feed))

```

PAST FEEDS

Bring in the relevant model outputs.

```{r}

this_feed <- "past"

biomass_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "biomass_", full.names = TRUE)

faeces_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "excretion_", full.names = TRUE)

feed_waste_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "feed_waste_", full.names = TRUE)

ammonia_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "NH4_", full.names = TRUE)

weight_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "weight_", full.names = TRUE)

```

Past feeds - loop through the locations file to add the biomass, excretion, faeces, uneaten feed in N.

```{r}

feed_type <-  "past"

commercial_size <- read.csv(sprintf(here("data/%s/params/Parameters_%s.csv"), this_species, feed_type)) %>% filter(Description == "Commercial size") %>% pull(Value) %>% as.numeric()

farm_ids <- locations_file$farm_id

this_id <- farm_ids[[29]]

#Set up multicore

future::plan(strategy = "multisession", workers = parallel::detectCores()-2)

farm_N_past <-

  future_map(.x = farm_ids, .f = \(this_id){

  message("Synthesising nitrogen data at harvest this farmID_", this_id)

  weight_days <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(days)
  weights <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(upper_bound)

  harvest_day <- 
    weight_days %>% 
    pluck(if(tail(weights, n=2)[1]>commercial_size){
      head(which(weights>commercial_size), n =2)[1]
    } else {
      tail(weight_days, n=2)[1]
      } )

   #biomass
  this_biomass_mean <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(biomass)/1e+6

  this_biomass_lb <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(lower_bound)/1e+6

   this_biomass_ub <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(upper_bound)/1e+6

   #weight
   this_weight_mean <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(weight)

   this_weight_lb <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(lower_bound)

   this_weight_ub <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day)%>% pull(upper_bound)

  #faeces 
    this_faeces_mean <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_excretion) %>% sum()/6.25/1e+6

     this_faeces_lb <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

  this_faeces_ub <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

  #feed waste
  this_feed_waste_mean <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_waste) %>% sum()/6.25/1e+6

  this_feed_waste_lb <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

  this_feed_waste_ub <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

  #excretion (NH4)
  this_ammonia_mean <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(nh4_production) %>% sum()/1e+6

  this_ammonia_lb <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/1e+6

  this_ammonia_ub <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/1e+6

  return(data.frame(farm_id = this_id, biomass_mean = this_biomass_mean, biomass_lb = this_biomass_lb, biomass_ub = this_biomass_ub, faeces_mean = this_faeces_mean, faeces_lb = this_faeces_lb, faeces_ub = this_faeces_ub, excretion_mean = this_ammonia_mean, excretion_lb = this_ammonia_lb, excretion_ub = this_ammonia_ub, feed_waste_mean = this_feed_waste_mean, feed_waste_lb = this_feed_waste_lb, feed_waste_ub = this_feed_waste_ub, total_N_mean = sum(this_faeces_mean, this_ammonia_mean, this_feed_waste_mean), total_N_lb = sum(this_faeces_lb, this_ammonia_lb, this_feed_waste_lb), total_N_ub = this_feed_waste_ub))

}) %>% bind_rows()

farm_N_past_spatial <- 
  locations_file %>% 
  select(iso3c, country, tonnes_per_farm, geometry, farm_id) %>% 
  distinct() %>% 
  left_join(farm_N_past)

qsave(x = farm_N_past_spatial, file =  sprintf(here("data/atlantic_salmon/data_products/model_summaries/%s/nitrogen_summary.qs"), this_feed))

```

FUTURE FEEDS

Bring in the relevant model outputs.

```{r}

this_feed <- "future"

biomass_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "biomass_", full.names = TRUE)

faeces_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "excretion_", full.names = TRUE)

feed_waste_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "feed_waste_", full.names = TRUE)

ammonia_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "NH4_", full.names = TRUE)

weight_files <- list.files(path = sprintf(here("data/atlantic_salmon/data_products/model_outputs/%s"), this_feed), pattern = "weight_", full.names = TRUE)

```

Future feeds - loop through the locations file to add the biomass, excretion, faeces, uneaten feed in N.

```{r}

feed_type <-  "future"

commercial_size <- read.csv(sprintf(here("data/%s/params/Parameters_%s.csv"), this_species, feed_type)) %>% filter(Description == "Commercial size") %>% pull(Value) %>% as.numeric()

farm_ids <- locations_file$farm_id

this_id <- farm_ids[[29]]

#Set up multicore

future::plan(strategy = "multisession", workers = parallel::detectCores()-2)

farm_N_future <-

  future_map(.x = farm_ids, .f = \(this_id){

  message("Synthesising nitrogen data at harvest this farmID_", this_id)

  weight_days <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(days)
  weights <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% pull(upper_bound)

  harvest_day <- 
    weight_days %>% 
    pluck(if(tail(weights, n=2)[1]>commercial_size){
      head(which(weights>commercial_size), n =2)[1]
    } else {
      tail(weight_days, n=2)[1]
      } )

    #biomass
  this_biomass_mean <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(biomass)/1e+6

  this_biomass_lb <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(lower_bound)/1e+6

   this_biomass_ub <- qread(biomass_files[grep(pattern = sprintf("farmID_%s.qs", this_id), biomass_files)]) %>% 
    slice(harvest_day) %>% pull(upper_bound)/1e+6

   #weight
   this_weight_mean <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(weight)

   this_weight_lb <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day) %>% pull(lower_bound)

   this_weight_ub <- qread(weight_files[grep(pattern = sprintf("farmID_%s.qs", this_id), weight_files)]) %>% filter(days == harvest_day)%>% pull(upper_bound)

  #faeces 
    this_faeces_mean <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_excretion) %>% sum()/6.25/1e+6

     this_faeces_lb <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

  this_faeces_ub <- qread(faeces_files[grep(pattern = sprintf("farmID_%s.qs", this_id), faeces_files)]) %>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

  #feed waste
  this_feed_waste_mean <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(mean_waste) %>% sum()/6.25/1e+6

  this_feed_waste_lb <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/6.25/1e+6

  this_feed_waste_ub <- qread(feed_waste_files[grep(pattern = sprintf("farmID_%s.qs", this_id), feed_waste_files)])%>% filter(nutrient == "Protein") %>% slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/6.25/1e+6

  #excretion (NH4)
  this_ammonia_mean <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(nh4_production) %>% sum()/1e+6

  this_ammonia_lb <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(lower_bound) %>% sum()/1e+6

  this_ammonia_ub <- qread(ammonia_files[grep(pattern = sprintf("farmID_%s.qs", this_id), ammonia_files)]) %>% 
    slice(1:harvest_day) %>% pull(upper_bound) %>% sum()/1e+6

  return(data.frame(farm_id = this_id, biomass_mean = this_biomass_mean, biomass_lb = this_biomass_lb, biomass_ub = this_biomass_ub, faeces_mean = this_faeces_mean, faeces_lb = this_faeces_lb, faeces_ub = this_faeces_ub, excretion_mean = this_ammonia_mean, excretion_lb = this_ammonia_lb, excretion_ub = this_ammonia_ub, feed_waste_mean = this_feed_waste_mean, feed_waste_lb = this_feed_waste_lb, feed_waste_ub = this_feed_waste_ub, total_N_mean = sum(this_faeces_mean, this_ammonia_mean, this_feed_waste_mean), total_N_lb = sum(this_faeces_lb, this_ammonia_lb, this_feed_waste_lb), total_N_ub = this_feed_waste_ub))

}) %>% bind_rows()

farm_N_future_spatial <- 
  locations_file %>% 
  select(iso3c, country, tonnes_per_farm, geometry, farm_id) %>% 
  distinct() %>% 
  left_join(farm_N_future)

qsave(x = farm_N_future_spatial, file =  sprintf(here("data/atlantic_salmon/data_products/model_summaries/%s/nitrogen_summary.qs"), this_feed))

```