### 2025-12-11 docs-working/05_formulating_feeds.qmd

---
title: "Formulating experimental feeds"
author: Tormey Reimer
date: today
editor: source
---

# Introduction

The purpose of this markdown is to formulate all the feeds that will be used in the analysis. These feeds are:
- marine_dominant: a "past" salmonid diet where most protein is sourced from marine-based ingredients.
- plant_dominant: a "current" salmonid diet where most protein is sourced from land-based ingredients.
- novel_inclusive_2: a potential future salmonid diet which includes some novel ingredients. This formulation uses maximal digestibility values.
- novel_inclusive_1: a potential future salmonid diet which includes some novel ingredients. This formulation uses the same formulation as novel_inclusive_2, but with minimal digestibility values.

```{r setup, include=F, cache=F}
#| code-summary: R setup

library(arrow)
library(sf)
library(dplyr)
library(tidyr)
library(terra)
library(magrittr)
library(purrr)
library(furrr)
library(future)
library(tictoc)
library(ggplot2)
library(fs)
library(conflicted)
library(stringr)
library(readxl)
library(units)
library(qs)
library(here)
conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% walk(source)
```

# Import ingredients

There is no check or correction here to ensure that protein + lipid + ash + carb = 1 in the incoming ingredient data. This was checked manually prior to import.

``` {r import-ingredients}
#| code-summary: Get ingredient details (PLC composition, digestibility)

ingreds <- file.path(input_feed_profile_path, "all_ingredients.xlsx") %>% 
  read_excel(sheet = "all_ingredients") %>% 
  mutate(ingredient = as.factor(ingredient))

ingred_nms <- levels(ingreds$ingredient)
```

# Import feeds

Unlike ingredients, the feeds do not need to be checked for sum = 1 prior to import. That check is done here.

``` {r import-feeds}
#| code-summary: Feed compositions

feed_inputs <- file.path(input_feed_profile_path, "all_feeds.xlsx") %>% 
  read_excel(sheet = "all_feeds")

feed_inputs <- feed_inputs %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "category"))) %>% 
  mutate(
    feed = as.factor(feed),
    ingredient = as.factor(ingredient),
    proportion = case_when(is.na(proportion) ~ 0, T ~ proportion)
  ) %>% 
  left_join(ingreds, by = "ingredient")
```

Include the range of digestibility values so that feeds can be sampled (if required) within the targets pipeline.

```{r min-max-values}
#| code-summary: Novel feeds use the mean digestibility values

feed_inputs <- feed_inputs %>% 
  rename(
    protein_digestibility = mean_protein_digestibility,
    carb_digestibility = mean_carb_digestibility,
    lipid_digestibility = mean_lipid_digestibility
  )

feed_types <- levels(feed_inputs$feed)
```

Check that ingredient proportions sum to 1.

```{r check-proportions}
#| code-summary: Composition proportions

feed_inputs <- feed_inputs %>% 
  filter(!is.na(proportion) & proportion != 0) %>% 
  group_by(feed) %>% 
  mutate(
    sum = sum(proportion),
    proportion = proportion/sum
  ) %>% 
  ungroup() %>% 
  select(-sum, -ash)
```

# Prepare for pipeline

The following chunk allows as many ingredients and feeds to be added to the incoming data as needed, it will still be formatted into a form that the `targets` pipelines can use.

```{r formulate-feeds}
#| code-summary: Format feeds as lists to feed into targets

feed_types <- levels(feed_inputs$feed)

feed_params <- purrr::map(feed_types, function(ft) {
  # Filter by specified feed
  df <- feed_inputs %>% filter(feed == ft)

  list(
    Proteins = df %>% 
      select(ingredient, proportion, contains("protein")) %>% 
      rename(
        macro = protein, 
        digest = protein_digestibility,
        min_digest = min_protein_digestibility,
        max_digest = max_protein_digestibility
      ),
    Carbohydrates = df %>% 
      select(ingredient, proportion, contains("carb")) %>% 
      rename(
        macro = carb, 
        digest = carb_digestibility,
        min_digest = min_carb_digestibility,
        max_digest = max_carb_digestibility
      ),
    Lipids = df %>% 
      select(ingredient, proportion, contains("lipid")) %>% 
      rename(
        macro = lipid, 
        digest = lipid_digestibility,
        min_digest = min_lipid_digestibility,
        max_digest = max_lipid_digestibility
      )
  )
}) %>% 
  setNames(feed_types)
```

```{r save}
#| code-summary: Save feed params file

qsave(feed_params, file.path(output_species_data_path, "feed_params.qs"))
```